---
title: "Portfolio Substitution between Coastal Pelagic Species under Shifting Target Species Distributions and Policy Constraints"
author:
  - name: Felipe Quezada
    email: felipe.quezada@noaa.gov
    institute: [ucsc, noaa1]
    correspondence: true
  - name: Desiree Tommasi
    institute: [ucsc, noaa1]
  - name: Stephen Stohs
    institute: noaa1
  - name: Timothy Frawley
    institute: [ucsc, noaa1]
  - name: Jonathan Sweeney 
    institute: noaa3
  - name: Isaac Kaplan
    institute: noaa2
  - name: Barbara Muhling
    institute: [ucsc, noaa1]  

institute:
  - ucsc: Institute of Marine Sciences, University of California Santa Cruz
  - noaa1: NOAA Southwest Fisheries Science Center
  - noaa2: NOAA Northwest Fisheries Science Center
  - noaa3: NOAA Pacific Islands Fisheries Science Center
  
date: "`r format(Sys.time(), '%B %e, %Y')`"
abstract: "Fishers often target a diverse set of species. These fishing portfolios allow them to reduce income risk. More diverse portfolios may increase fisher resilience to climate-driven changes in target species’ spatial distributions and availability. Therefore, evaluations of climate risk to a particular fishery need to consider climate impacts on the availability of multiple target species and how such changes may impact substitution behaviour. Regulations and other constraints (e.g., port infrastructure constraints on where landings of a particular species may occur, or permit requirements) may also reduce the degree of substitution we observe. However, what factors influence diversification and how climate driven changes in target species distribution may interact with other constraints to affect changes in substitution behaviour remains unclear for many fisheries. In this study, we analyze how historical changes in forage species distribution and the closure of the Pacific sardine fishery affected landing substitution between two coastal pelagic species: Pacific sardine (*Sardinops sagax*) and market squid (*Doryteuthis opalescens*), that are targeted by the US West Coast Coastal Pelagic Fleet. Using a discrete choice modeling approach embedded in our landing model, we also study how spatial distribution and closure affected the coastal pelagic species fisheries' participation decisions over the 2000-2019 period. Our preliminary results show strong substitution between market squid and Pacific sardine when both were available, while the Pacific sardine closure in 2015 was associated with reduced market squid landings suggesting lower fishers’ participation in this fishery."
header-includes:
  \usepackage{mathtools}
  \usepackage{tcolorbox}
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{times}
  \usepackage{setspace}
  \usepackage{rotating}
  \clearpage
  \setstretch{1.5}
output:
  pdf_document:
    number_sections: yes
    citation_package: natbib
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: references.bib
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
rm(list = ls(all.names = TRUE)) 
gc()
# load("stan_fit.RData")

library(bookdown)
library(brms)
library(cluster)
library(cmdstanr)
library(data.table)
library(distances)
library(doBy)
library(dplyr)
library(fastDummies)
library(forcats)
library(ggplot2)
library(here)
library(hrbrthemes)
library(kableExtra)
library(lmtest)
library(lubridate)
library(magrittr)
library(Rcpp)
library(patchwork)
library(plm)
library(papaja)
library(reshape)
library(reshape2)
library(rstan)
library(scales)
library(sjlabelled)
library(texreg)
library(tidyr)
library(tidyverse)
library(tinytex)
library(viridis)
library(xtable)
library(zoo)

```

```{r read_data, include=FALSE}
port_area <- read.csv(file = here::here("Data", "Ports", "ports_area_and_name_codes.csv"))
  PacFIN.month <- read.csv(file ="C:\\Data\\PacFIN data\\PacFIN_month.csv")
  
  ### Include port area code (using PacFIN Port Code)
  PacFIN.month <- PacFIN.month %>% merge(port_area, by = c("PACFIN_PORT_CODE"), all.x = TRUE)
  PacFIN.month.CPS <- PacFIN.month %>% 
    dplyr::filter(PACFIN_SPECIES_CODE %in% 
    c("CMCK", "JMCK", "MSQD", "NANC", "PSDN"))
    
  rm(port_area)
  
```


```{r functions, include=FALSE}
meanfun <- function(x, ...){
  c(mean=mean(x, na.rm=TRUE, ...)) #, v=var(x, na.rm=TRUE, ...), l=length(x))
}
sumfun <- function(x, ...){
  c(sum=sum(x, na.rm=TRUE, ...)) #, v=var(x, na.rm=TRUE, ...), l=length(x))
}
sum_mean_fun <- function(x, ...){
  c(mean=mean(x, na.rm=TRUE, ...), sum=sum(x, na.rm=TRUE, ...)) #, l=length(x))
}

```

# Descriptive figures 

```{r psdn_msqd_month_by_port, echo=FALSE, fig.cap="Fishing season for Pacific sardine and Market squid (average monthly total harvest for the period 2000-2014, before the Pacific sardine closure).\\label{fig:fish_season}", message=FALSE, warning=FALSE}

port_names <- as_labeller(c("SP" = "San Pedro", "HNM" = "Hueneme", "MNT" = "Monterrey", "MOS" = "Moss Landing", 
                               "LWC" = "Ilwaco / Chinook", "AST" = "Astoria", 
                               "PRN" = "Princeton/ Half Moon\nBay", "SF" = "San Francisco", 
                               "SLT" = "Sausalito", "TRM" = "Terminal Island",
                               "VEN" = "Ventura", "WPT" = "Westport"))

q.psdn.by.month <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>%  
  group_by(LANDING_YEAR, LANDING_MONTH, PACFIN_PORT_CODE) %>% 
  summarise(Landings_PSDN = sum(LANDED_WEIGHT_MTONS.sum)) %>%  
  group_by(LANDING_MONTH, PACFIN_PORT_CODE) %>% 
  summarise(Landings_PSDN = mean(Landings_PSDN)) %>%
    filter(PACFIN_PORT_CODE == "AST" | PACFIN_PORT_CODE == "SP" |
             PACFIN_PORT_CODE == "TRM" | PACFIN_PORT_CODE == "MOS" | 
             PACFIN_PORT_CODE == "WPT") 
  
g1 <- ggplot(data=q.psdn.by.month, aes(x=LANDING_MONTH, y=Landings_PSDN)) +
    geom_bar(stat="identity", fill=rgb(0.1,0.4,0.5,0.7), width=0.4) + 
  scale_x_continuous(name = "Month", breaks=1:12) +
  scale_y_continuous(name = "Landings (M tons)" ) + 
  facet_grid(~ factor(PACFIN_PORT_CODE, levels=c("SP", "TRM", "MOS", "AST", "WPT")), 
             labeller = port_names) + ggtitle("(a) Pacific sardine") + 
  theme(plot.title = element_text(size=9, face="bold.italic"), axis.text = element_text(size = 6))

q.msqd.by.month <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>%  
  group_by(LANDING_YEAR, LANDING_MONTH, PACFIN_PORT_CODE) %>% 
  summarise(Landings_PSDN = sum(LANDED_WEIGHT_MTONS.sum)) %>%  
  group_by(LANDING_MONTH, PACFIN_PORT_CODE) %>% 
  summarise(Landings_PSDN = mean(Landings_PSDN)) %>%
    filter(PACFIN_PORT_CODE == "SP" | PACFIN_PORT_CODE == "VEN" | PACFIN_PORT_CODE == "TRM" | 
              PACFIN_PORT_CODE == "MOS" | PACFIN_PORT_CODE == "MNT") 

g2 <- ggplot(data=q.msqd.by.month, aes(x=LANDING_MONTH, y=Landings_PSDN)) +
    geom_bar(stat="identity", fill=rgb(0.1,0.4,0.5,0.7), width=0.4) + 
  scale_x_continuous(name = "Month", breaks=1:12) +
  scale_y_continuous(name = "Landings (M tons)" ) + 
  facet_grid(~factor(PACFIN_PORT_CODE, levels = c("SP", "TRM", "VEN", "MNT", "MOS")), 
             labeller = port_names) + ggtitle("(b) Market squid") + 
  theme(plot.title = element_text(size=9, face="bold.italic"), axis.text = element_text(size = 6))

g1 / g2

rm(g1, g2, q.psdn.by.month, q.msqd.by.month, port_names)
```

```{r avg_land_rev_price, echo=FALSE, fig.cap="Average annual landings, prices and revenues by species (1981-2014).\\label{fig:avg_lan_rev}"}
## Landings chart
landings.year <- summaryBy(LANDED_WEIGHT_MTONS.sum ~ PACFIN_SPECIES_CODE + LANDING_YEAR, FUN=sumfun, data=PacFIN.month.CPS)
  landings.year <- landings.year %>% filter(LANDING_YEAR <2015)
  landings.avg.year <- summaryBy(LANDED_WEIGHT_MTONS.sum.sum ~ PACFIN_SPECIES_CODE, FUN=meanfun, da=landings.year)

g1 <- ggplot(landings.avg.year, aes(PACFIN_SPECIES_CODE, LANDED_WEIGHT_MTONS.sum.sum.mean)) +
  geom_col(width=0.4) + ggtitle("(a) Annual landings") + 
  theme(plot.title = element_text(size=9, face="bold.italic"), axis.text = element_text(size = 7), 
        axis.title = element_text(size = 8)) +
  xlab("") + ylab("Landings (tons)") + 
  scale_x_discrete(labels=c("CMCK" = "Pacific\nMackerel", "JMCK" = "Jack\nMackerel",
                            "MSQD" = "Market\nSquid", "NANC" = "Northern\nAnchovy",
                            "PSDN" = "Pacific\nSardine"))

## Landings chart
price.year <- summaryBy(AFI_PRICE_PER_MTON.mean ~ PACFIN_SPECIES_CODE + LANDING_YEAR, FUN=meanfun, da=PacFIN.month.CPS)
  price.year <- price.year %>% filter(LANDING_YEAR <2015)
  price.avg.year <- summaryBy(AFI_PRICE_PER_MTON.mean.mean ~ PACFIN_SPECIES_CODE, FUN=meanfun, da=price.year)

g2 <- ggplot(price.avg.year, aes(PACFIN_SPECIES_CODE, AFI_PRICE_PER_MTON.mean.mean.mean)) +
  geom_col(width=0.2) + ggtitle("(c) Average annual price") + 
  theme(plot.title = element_text(size=9, face="bold.italic"), axis.text = element_text(size = 7), 
        axis.title = element_text(size = 8)) +
  xlab("") + ylab("Price (USD/ton)") + 
  scale_x_discrete(labels=c("CMCK" = "Pacific\nMackerel", "JMCK" = "Jack\nMackerel",
                            "MSQD" = "Market\nSquid", "NANC" = "Northern\nAnchovy",
                            "PSDN" = "Pacific\nSardine"))

## Revenue chart

revenue.year <- summaryBy(AFI_EXVESSEL_REVENUE.sum ~ PACFIN_SPECIES_CODE + LANDING_YEAR, FUN=sumfun, data=PacFIN.month.CPS)
  revenue.year <- revenue.year %>% filter(LANDING_YEAR <2015)
  revenue.avg.year <- summaryBy(AFI_EXVESSEL_REVENUE.sum.sum ~ PACFIN_SPECIES_CODE, FUN=meanfun, data=revenue.year)

g3 <- ggplot(revenue.avg.year, aes(PACFIN_SPECIES_CODE, AFI_EXVESSEL_REVENUE.sum.sum.mean)) +
  geom_col(width=0.4) + ggtitle("(b) Annual revenue") + 
  theme(plot.title = element_text(size=9, face="bold.italic"), axis.text = element_text(size = 7), 
        axis.title = element_text(size = 8)) +
  xlab("") + ylab("Annual revenues (Millions of USD)") +  scale_y_continuous(labels = label_number(scale = 1e-6)) +
  scale_x_discrete(labels=c("CMCK" = "Pacific\nMackerel", "JMCK" = "Jack\nMackerel",
                            "MSQD" = "Market\nSquid", "NANC" = "Northern\nAnchovy",
                            "PSDN" = "Pacific\nSardine"))

(g1 + g3) / g2

rm(g1, g2, g3, landings.year, landings.avg.year, revenue.year, revenue.avg.year, price.year, price.avg.year)

```

```{r datasets_year, include=FALSE}
# Calculate number of vessels by species per month # 
nvessel.year <- PacFIN.month.CPS %>% 
  dplyr::select(PACFIN_SPECIES_CODE, LANDING_YEAR, VESSEL_NUM) %>% unique() %>% 
  mutate(n_vessel = 1) %>% group_by(PACFIN_SPECIES_CODE, LANDING_YEAR) %>%
  summarise(n_vessel = sum(n_vessel))

# Calculate average price and total landings by species per month #
landing.price.year <- summaryBy(LANDED_WEIGHT_MTONS.sum + AFI_PRICE_PER_MTON.mean ~ 
                                  PACFIN_SPECIES_CODE + LANDING_YEAR, 
                      FUN=sum_mean_fun, data=PacFIN.month.CPS) 

```

```{r avg_landings_by_port, echo=FALSE, fig.cap="Annual average landings by port area.\\label{fig:avg_landings_by_ports}", message=FALSE, warning=FALSE}
landings.by.port.year <- summaryBy(LANDED_WEIGHT_MTONS.sum ~ 
  PACFIN_SPECIES_CODE + PACFIN_PORT_CODE + LANDING_YEAR + AGENCY_CODE, FUN=sumfun, data=PacFIN.month.CPS)
  landings.by.port.avg.year <- summaryBy(LANDED_WEIGHT_MTONS.sum.sum ~ PACFIN_SPECIES_CODE + PACFIN_PORT_CODE 
    + AGENCY_CODE, FUN=meanfun, data=landings.by.port.year) %>% 
    filter(PACFIN_SPECIES_CODE == "PSDN" | PACFIN_SPECIES_CODE == "CMCK" | 
           PACFIN_SPECIES_CODE == "MSQD" |  PACFIN_SPECIES_CODE == "NANC") %>%
    filter(PACFIN_PORT_CODE == "AST" | PACFIN_PORT_CODE == "HNM" | PACFIN_PORT_CODE == "SP" |
           PACFIN_PORT_CODE == "VEN" | PACFIN_PORT_CODE == "TRM" | PACFIN_PORT_CODE == "MOS" |
           PACFIN_PORT_CODE == "WPT" | PACFIN_PORT_CODE == "MNT" | PACFIN_PORT_CODE == "LWC") %>%
    mutate(PACFIN_PORT_CODE = fct_relevel(PACFIN_PORT_CODE, 
      "SP", "WLM", "TRM", "HNM", "VEN", "MNT", "MOS", "PRN", "SF", "SLT", "WIN", "AST", "LWC", "WPT"))

 # PACFIN_PORT_CODE == "PRN" | PACFIN_PORT_CODE == "WLM" | PACFIN_PORT_CODE == "WIN"
 
states_names <- as_labeller(c(`C` = "California", `O` = "Oregon",`W` = "Washington"))
  ggplot(landings.by.port.avg.year, aes(fill=PACFIN_SPECIES_CODE, 
                                      y=LANDED_WEIGHT_MTONS.sum.sum.mean, x=PACFIN_PORT_CODE)) +
  geom_bar(position="dodge", stat="identity") + 
  facet_grid(~ AGENCY_CODE, scale="free", space="free_x", labeller = states_names) + 
  theme(strip.text.x = element_text(size = 7)) +
  theme(legend.position="bottom") + 
  scale_fill_viridis(discrete = T, labels=c("MSQD" = "Market Squid", "NANC" = "Northern Anchovy", 
                               "PHRG" = "Pacific Herring", "PSDN" = "Pacific Sardine", 
                               "BTNA" = "Bluefin tuna", "JMCK" = "Jack mackerel", 
                               "CMCK" = "Pacific mackerel", "UMCK" = "Unspecified mackerel",
                               "PBNT" = "Pacific bonito")) +
  theme(axis.text.x = element_text(angle=90, hjust=0.95, vjust=0.45)) + xlab("Ports") + 
  ylab("Landings (tons)") + guides(fill=guide_legend(title="Species: "))  + 
    scale_x_discrete(labels=c("SP" = "San\nPedro", "HNM" = "Hueneme", 
                              "MNT" = "Monterrey", "MOS" = "Moss\nLanding", 
                              "LWC" = "Ilwaco /\nChinook", "AST" = "Astoria", 
                              "PRN" = "Princeton/\nHalf Moon\nBay", "TRM" = "Terminal\nIsland",
                              "VEN" = "Ventura", "WIN" = "Winchester\nBay",
                              "WLM" = "Willmington", "WPT" = "Westport")) +
    scale_color_brewer(palette="Set2")

  
  
rm(states_names, landings.by.port.avg.year)
```

```{r ts_landing_price, echo=FALSE, fig.cap="Evolution of annual total landings and average annual prices by CPS species during the period 1981-2020.\\label{fig:ts_landings_price}", message=FALSE}
landing.price.year.sel <- landing.price.year %>% 
  dplyr::select("LANDED_WEIGHT_MTONS.sum.sum", "AFI_PRICE_PER_MTON.mean.mean",
                "PACFIN_SPECIES_CODE", "LANDING_YEAR")

# Graph landing v/s number of  vessels
coeff1 <- 200
  df <- landing.price.year.sel %>% filter(PACFIN_SPECIES_CODE == "MSQD")
  df$AFI_PRICE_PER_MTON.mean.mean <- df$AFI_PRICE_PER_MTON.mean.mean * coeff1 
  df <- gather(df, key = Variable, value = value,
               c("LANDED_WEIGHT_MTONS.sum.sum", "AFI_PRICE_PER_MTON.mean.mean"))
g1 <- ggplot(df, aes(x=LANDING_YEAR, y = value, colour = Variable)) + geom_line(size=1) +
    scale_y_continuous(name = "Landings (M Tons)", sec.axis = sec_axis(~./coeff1, name = "Price (USD/Ton)")) +
    theme(legend.position="none") + scale_x_continuous(name = element_blank()) + 
    theme(plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
      ggtitle("(b) Market Squid")  +  
    scale_color_brewer(palette="Set2",
                       limits = c("LANDED_WEIGHT_MTONS.sum.sum", "AFI_PRICE_PER_MTON.mean.mean"))

coeff2 <- 100
df <- landing.price.year.sel %>% filter(PACFIN_SPECIES_CODE == "PSDN")
  df$AFI_PRICE_PER_MTON.mean.mean <- df$AFI_PRICE_PER_MTON.mean.mean * coeff2 
  df <- gather(df, key = Variable, value = value,
               c("LANDED_WEIGHT_MTONS.sum.sum", "AFI_PRICE_PER_MTON.mean.mean"))
  g2 <- ggplot(df, aes(x=LANDING_YEAR, y = value, colour = Variable)) + geom_line(size=1) +
    scale_y_continuous(name = "Landings (M Tons)", sec.axis = sec_axis(~./coeff2, name = "Price (USD/Ton)")) + 
    theme(legend.position="right") + scale_x_continuous(name = "Year") +
    theme(plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
      ggtitle("(a) Pacific Sardine")  + guides(colour=guide_legend(title="Variables: ")) + 
  scale_color_brewer(palette="Set2",
                     limits = c("LANDED_WEIGHT_MTONS.sum.sum", "AFI_PRICE_PER_MTON.mean.mean"),
                     labels=c("AFI_PRICE_PER_MTON.mean.mean" = "Price", "LANDED_WEIGHT_MTONS.sum.sum" = "Landings"))  

  (g2 / g1)
  rm(df, g1, g2, landing.price.year.sel)
  
  
```

```{r ts_Q_P_n, eval=FALSE, fig.cap="Yearly total landing, warning=FALSE, average price and total number of vessel by CPS species (1981 - 2020).\\label{fig:ts_landings}", message=FALSE, include=FALSE}

g1 <- ggplot(landing.price.year,
             aes(x = LANDING_YEAR, y = LANDED_WEIGHT_MTONS.sum.sum, group = PACFIN_SPECIES_CODE, colour = PACFIN_SPECIES_CODE)) +
  geom_line(size=1) + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Landings (MTons)") +
  theme(legend.position = "right", plot.title = element_text(size=9, face="bold.italic"),
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
  ggtitle("(a) Landings") +   guides(colour=guide_legend(title="Species: ")) +
  scale_color_brewer(palette="Set2", labels=c("MSQD" = "Market Squid",
                                              "NANC" = "Northern Anchovy",
                                              "PSDN" = "Pacific Herring",
                                              "PSDN" = "Pacific Sardine"))

g2 <- ggplot(landing.price.year,
             aes(x = LANDING_YEAR, y = AFI_PRICE_PER_KG.mean.mean, group = PACFIN_SPECIES_CODE, colour = PACFIN_SPECIES_CODE)) +
  geom_line(size=1) + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Price ($/Kg)")  +
   scale_color_brewer(palette="Set2") +  theme(legend.position = "none", plot.title = element_text(size=9, face="bold.italic"),
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
  ggtitle("(b) Prices")


g3 <- ggplot(nvessel.year, aes(x = LANDING_YEAR, y = n_vessel, group = PACFIN_SPECIES_CODE, colour = PACFIN_SPECIES_CODE)) +
  geom_line(size=1) + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "# of vessels") +   theme(legend.position="none")  +
   scale_color_brewer(palette="Set2") +  theme(plot.title = element_text(size=9, face="bold.italic"),
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
  ggtitle("(c) Number of vessels")

(g2 + g1) / g3

rm(g1, g2, g3)

```

```{r SDM_land, eval=FALSE, fig.cap="Relationship between the probability of presence (from SDMs) and landings by CPS species in three main ports: San Pedro, Terminal Island, and Moss Landing.\\label{fig:sdm_land_by_port}", message=FALSE, warning=FALSE, include=FALSE}

sdm.by.species <- PacFIN.month.CPS %>%
  dplyr::select(LANDING_YEAR, PSDN_SDM_60, MSQD_SDM_90, NANC_SDM_20, LANDED_WEIGHT_MTONS.sum, PACFIN_SPECIES_CODE) %>% 
  filter(LANDING_YEAR >= 1998 & LANDING_YEAR <= 2019) %>%
  group_by(LANDING_YEAR, PACFIN_SPECIES_CODE) %>% 
  summarize(PSDN_SDM_60 = mean(PSDN_SDM_60, na.rm = TRUE), 
            NANC_SDM_20 = mean(NANC_SDM_20, na.rm = TRUE), 
            MSQD_SDM_90 = mean(MSQD_SDM_90, na.rm = TRUE),
            LANDED_WEIGHT_MTONS = sum(LANDED_WEIGHT_MTONS.sum, na.rm = TRUE)) %>%
  spread(PACFIN_SPECIES_CODE, LANDED_WEIGHT_MTONS) %>% 
  dplyr::rename(Landings_PSDN = PSDN) %>% dplyr::rename(Landings_MSQD = MSQD) %>% 
  dplyr::rename(Landings_NANC = NANC) %>%
  group_by(LANDING_YEAR) %>% 
  summarize(PSDN_SDM_60 = mean(PSDN_SDM_60, na.rm = TRUE), 
            NANC_SDM_20 = mean(NANC_SDM_20, na.rm = TRUE), 
            MSQD_SDM_90 = mean(MSQD_SDM_90, na.rm = TRUE),
            Landings_PSDN = sum(Landings_PSDN, na.rm = TRUE), Landings_MSQD = sum(Landings_MSQD, na.rm = TRUE),
            Landings_NANC = sum(Landings_NANC, na.rm = TRUE)) 
  # sdm.by.species$Date <- zoo::as.yearmon(paste(sdm.by.species$LANDING_YEAR, sdm.by.species$LANDING_MONTH, sep='-'))

# Plot
coeff <- 200000
g1 <-  ggplot(sdm.by.species) + 
  geom_line(mapping = aes(x = LANDING_YEAR, y = Landings_PSDN, color = "Landings"), size = 0.5) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = PSDN_SDM_60*coeff, color = "Probability of presence"), 
            size = 0.5, linetype = "dashed") + 
  scale_x_continuous(name = element_blank(), labels = NULL) +
  scale_y_continuous(name = element_blank(), sec.axis = sec_axis(~./coeff, name = "")) +
  theme(legend.position="right", plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
  ggtitle("(a) Pacific sardine") +  scale_color_manual(name = "Variable: ", 
                     values = c("Landings" = "grey", "Probability of presence" = "blue"))

coeff2 <- 300000
g2 <- ggplot(sdm.by.species) + 
  geom_line(mapping = aes(x = LANDING_YEAR, y = Landings_MSQD), size = 0.5, color = "grey") +
  geom_line(mapping = aes(x = LANDING_YEAR, y = MSQD_SDM_90*coeff2), 
            size = 0.5, color = "blue", linetype = "dashed") + 
  scale_x_continuous(name = element_blank(), labels = NULL) +
  scale_y_continuous(name = "Landings", sec.axis = sec_axis(~./coeff2, name = "P(presence)")) +
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
  ggtitle("(b) Market squid")

g1 / g2

rm(g1, g2, sdm.by.species, coeff, coeff2, coeff3)

```

```{r SDM_land_by_port, echo=FALSE, fig.cap="Landings v/s probability of presence by port area.\\label{fig:sdm_land_by_port}", message=FALSE, warning=FALSE}

sdm.by.species <- PacFIN.month.CPS %>%
  dplyr::select(LANDING_YEAR, PORT_NAME, PSDN_SDM_60, MSQD_SDM_90, MSQD_SPAWN_SDM_90, NANC_SDM_20, 
                LANDED_WEIGHT_MTONS.sum, PACFIN_SPECIES_CODE, PACFIN_PORT_CODE) %>% 
  filter(LANDING_YEAR >= 1998 & LANDING_YEAR <= 2019) %>%
  group_by(LANDING_YEAR, PORT_NAME, PACFIN_PORT_CODE, PACFIN_SPECIES_CODE) %>% 
  summarize(PSDN_SDM_60 = mean(PSDN_SDM_60, na.rm = TRUE), 
            NANC_SDM_20 = mean(NANC_SDM_20, na.rm = TRUE), 
            MSQD_SDM_90 = mean(MSQD_SDM_90, na.rm = TRUE),
            MSQD_SPAWN_SDM_90 = mean(MSQD_SPAWN_SDM_90, na.rm=TRUE),
            LANDED_WEIGHT_MTONS = sum(LANDED_WEIGHT_MTONS.sum, na.rm = TRUE)) %>%
  spread(PACFIN_SPECIES_CODE, LANDED_WEIGHT_MTONS) %>% 
  dplyr::rename(Landings_PSDN = PSDN) %>% dplyr::rename(Landings_MSQD = MSQD) %>% 
  dplyr::rename(Landings_NANC = NANC) %>%
  filter(PACFIN_PORT_CODE == "SP" | PACFIN_PORT_CODE == "TRM" | PACFIN_PORT_CODE == "MOS") %>% 
  mutate(PACFIN_PORT_CODE = fct_relevel(PACFIN_PORT_CODE, "SP", "TRM", "MOS")) %>%
  group_by(LANDING_YEAR, PORT_NAME, PACFIN_PORT_CODE) %>% 
  summarize(PSDN_SDM_60 = mean(PSDN_SDM_60, na.rm = TRUE), 
            NANC_SDM_20 = mean(NANC_SDM_20, na.rm = TRUE), 
            MSQD_SDM_90 = mean(MSQD_SDM_90, na.rm = TRUE),
            MSQD_SPAWN_SDM_90 = mean(MSQD_SPAWN_SDM_90, na.rm = TRUE),
            Landings_PSDN = sum(Landings_PSDN, na.rm = TRUE), 
            Landings_MSQD = sum(Landings_MSQD, na.rm = TRUE),
            Landings_NANC = sum(Landings_NANC, na.rm = TRUE))
  # sdm.by.species$Date <- zoo::as.yearmon(paste(sdm.by.species$LANDING_YEAR, sdm.by.species$LANDING_MONTH, sep='-'))

# Plot
coeff <- 60000
g1 <-  ggplot(sdm.by.species) + 
  geom_line(mapping = aes(x = LANDING_YEAR, y = Landings_PSDN), size = 0.5, color = "grey") +
  geom_line(mapping = aes(x = LANDING_YEAR, y = PSDN_SDM_60*coeff), 
            size = 0.5, color = "blue", linetype = "dashed") + 
  facet_wrap(~ PORT_NAME) + 
  scale_x_continuous(name = element_blank(), labels = NULL) +
  scale_y_continuous(name = "Landings", sec.axis = sec_axis(~./coeff2, name = "P(presence)")) +
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
  ggtitle("(a) Pacific sardine (within 60 km radious)")

coeff2 <- 300000
g2 <- ggplot(sdm.by.species) + 
  geom_line(mapping = aes(x = LANDING_YEAR, y = Landings_MSQD, color = "Landings"), size = 0.5) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = MSQD_SPAWN_SDM_90*coeff2, color = "Probability of presence"), 
            size = 0.5, linetype = "dashed") + 
  facet_wrap(~ PORT_NAME) + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Landings", sec.axis = sec_axis(~./coeff2, name = "P(presence)")) +
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8),
        legend.position="bottom") + 
  ggtitle("(b) Market squid (Spawning aggregation model within 90 km radious)") +  scale_color_manual(name = "Variable: ", 
                     values = c("Landings" = "grey", "Probability of presence" = "blue"))

g1 / g2

rm(g1, g2, g3, sdm.by.species, coeff, coeff2, coeff3)

```




```{r SDM_land_by_port, echo=FALSE, fig.cap="Landings v/s probability of presence by port area.\\label{fig:sdm_land_by_port}", message=FALSE, warning=FALSE}

sdm.by.species <- PacFIN.month.CPS %>%
  dplyr::select(LANDING_YEAR, PORT_NAME, PSDN_SDM_60, MSQD_SDM_90, MSQD_SPAWN_SDM_90, 
                MSQD_SPAWN_SDM_100, MSQD_SPAWN_SDM_200, MSQD_SPAWN_SDM_300, 
                MSQD_SPAWN_SDM_5_100, NANC_SDM_20, 
                LANDED_WEIGHT_MTONS.sum, PACFIN_SPECIES_CODE, PORT_AREA_CODE) %>% 
  filter(LANDING_YEAR >= 2000 & LANDING_YEAR <= 2018) %>%
  group_by(LANDING_YEAR, PORT_AREA_CODE, PACFIN_SPECIES_CODE) %>% 
  summarize(MSQD_SDM_90 = mean(MSQD_SDM_90, na.rm = TRUE),
            MSQD_SPAWN_SDM_90 = mean(MSQD_SPAWN_SDM_90, na.rm=TRUE),
            LANDED_WEIGHT_MTONS = sum(LANDED_WEIGHT_MTONS.sum, na.rm = TRUE)) %>%
  spread(PACFIN_SPECIES_CODE, LANDED_WEIGHT_MTONS) %>% 
  dplyr::rename(Landings_PSDN = PSDN) %>% dplyr::rename(Landings_MSQD = MSQD) %>% 
  dplyr::rename(Landings_NANC = NANC) %>%
  filter(PORT_AREA_CODE == "LAA" | PORT_AREA_CODE == "MNA" | PORT_AREA_CODE == "SBA") %>% 
  mutate(PORT_AREA_CODE = fct_relevel(PORT_AREA_CODE, "LAA", "MNA", "SBA")) %>%
  group_by(LANDING_YEAR, PORT_AREA_CODE) %>% 
  summarize(MSQD_SDM_90 = mean(MSQD_SDM_90, na.rm = TRUE),
            MSQD_SPAWN_SDM_90 = mean(MSQD_SPAWN_SDM_90, na.rm = TRUE),
            Landings_PSDN = sum(Landings_PSDN, na.rm = TRUE), 
            Landings_MSQD = sum(Landings_MSQD, na.rm = TRUE),
            Landings_NANC = sum(Landings_NANC, na.rm = TRUE))
  # sdm.by.species$Date <- zoo::as.yearmon(paste(sdm.by.species$LANDING_YEAR, sdm.by.species$LANDING_MONTH, sep='-'))

# Plot
coeff2 <- 400000
g2 <- ggplot(sdm.by.species) + 
  geom_line(mapping = aes(x = LANDING_YEAR, y = Landings_MSQD, color = "Landings"), size = 0.5) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = MSQD_SPAWN_SDM_90*coeff2, color = "Probability of presence"),
            size = 0.5, linetype = "dashed") + 
  facet_wrap(~ PORT_AREA_CODE) + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Landings", sec.axis = sec_axis(~./coeff2, name = "P(presence)")) +
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8),
        legend.position="none") + 
  ggtitle("(a) Market squid (spawning biomass)") +  scale_color_manual(name = "Variable: ", 
                     values = c("Landings" = "grey", "Probability of presence" = "blue"))

coeff3 <- 200000
g3 <- ggplot(sdm.by.species) + 
  geom_line(mapping = aes(x = LANDING_YEAR, y = Landings_MSQD, color = "Landings"), size = 0.5) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = MSQD_SDM_90*coeff3, 
                          color = "Probability of presence"),
            size = 0.5, linetype = "dashed") + 
  facet_wrap(~ PORT_AREA_CODE) + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Landings", sec.axis = sec_axis(~./coeff3, name = "P(presence)")) +
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8),
        legend.position="none") + 
  ggtitle("(b) Market squid") +  scale_color_manual(name = "Variable: ",
           values = c("Landings" = "grey", "Probability of presence" = "blue"))
   

g2 / g3
  




# rm(g1, g2, g3, sdm.by.species, coeff, coeff2, coeff3)

```


```{r SDM_land_NANC_in_MNA, eval=FALSE, fig.cap="Landings v/s probability of presence by port area.", message=FALSE, warning=FALSE, include=FALSE}

collapse <- PacFIN_yearly %>%
  filter(Species_code == "NANC") %>%
  dplyr::select(Landing_year, Port, NANC_SDM_20_sep_dec, Landings, Species_code) %>% 
  filter(Port == "MNA") %>% mutate(Port = fct_relevel(Port, "MNA")) %>% 
  spread(Species_code, Landings) %>% dplyr::rename(Landings_NANC = NANC) %>%
  filter(Landing_year >= 2000 & Landing_year <= 2018)
  
# Plot
coeff <- 50000
g3 <- ggplot(collapse) + 
  geom_line(mapping = aes(x = Landing_year, y = Landings_NANC, color = "Landings"),
            size = 0.5) +
  geom_line(mapping = aes(x = Landing_year, y = NANC_SDM_20_sep_dec*coeff, color = "Probability of presence"), 
            size = 0.5, linetype = "dashed") + 
  facet_wrap(~ Port) + 
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = element_blank(), sec.axis = sec_axis(~./coeff, name = element_blank())) +
  theme(legend.position="bottom", plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
  ggtitle("Northern anchovy") +  scale_color_manual(name = "Variable: ", 
                     values = c("Landings" = "grey", "Probability of presence" = "blue"))

g3


collapse2 <- as.data.frame(diff(collapse$NANC_SDM_20_sep_dec))
  names(collapse2)[1] <- 'NANC_SDM_20_sep_dec'
  collapse2 <- collapse2 %>% mutate(Landings_NANC = diff(collapse$Landings_NANC)) %>% mutate(Port = "MNA") %>%
    mutate(Landing_year = seq(2001,2018))

dg3 <- ggplot(collapse2) + 
  geom_line(mapping = aes(x = Landing_year, y = Landings_NANC, color = "diff(Landings)"),
            size = 0.5) +
  geom_line(mapping = aes(x = Landing_year, y = NANC_SDM_20_sep_dec*coeff, color = "diff(Probability of presence)"), 
            size = 0.5, linetype = "dashed") + 
  facet_wrap(~ Port) + 
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = element_blank(), sec.axis = sec_axis(~./coeff, name = element_blank())) +
  theme(legend.position="bottom", plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
  ggtitle("Northern anchovy (first differences)") +  scale_color_manual(name = "Variable: ", 
                     values = c("diff(Landings)" = "grey", "diff(Probability of presence)" = "blue"))

dg3


library("writexl")
write_xlsx(collapse,"Data\\monterrey_NANC_sdm_landings.xlsx")

```

```{r ts_landings_vessels, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Evolution of Pacific sardine landings and the number of vessels landing market squid during the period 1981-2020.\\label{fig:ts_landings}"}
landing.nvessels.year.sel <- landing.price.year %>%
  inner_join(nvessel.year, by = c("PACFIN_SPECIES_CODE", "LANDING_YEAR")) %>%
  dplyr::select("LANDED_WEIGHT_MTONS.sum.sum", "n_vessel", "PACFIN_SPECIES_CODE", "LANDING_YEAR")

# Graph landing v/s number of  vessels
coeff <- 1500
  df <- landing.nvessels.year.sel %>% filter(PACFIN_SPECIES_CODE == "MSQD" | PACFIN_SPECIES_CODE == "PSDN") %>%
    mutate(n_vessel = ifelse(PACFIN_SPECIES_CODE == "MSQD", n_vessel, 0)) %>%
    mutate(LANDED_WEIGHT_MTONS.sum.sum = ifelse(PACFIN_SPECIES_CODE == "PSDN",
                                                LANDED_WEIGHT_MTONS.sum.sum, 0)) %>%
    group_by(LANDING_YEAR) %>%
    summarise(n_vessel = sum(n_vessel), LANDED_WEIGHT_MTONS.sum.sum = sum(LANDED_WEIGHT_MTONS.sum.sum))

  df$n_vessel <- df$n_vessel * coeff
  df <- gather(df, key = Variable, value = value,
               c("n_vessel", "LANDED_WEIGHT_MTONS.sum.sum"))
  ggplot(df, aes(x=LANDING_YEAR, y = value, colour = Variable)) + geom_line(size=1) +
    scale_y_continuous(name = "Landings of Pacific sardine (M Tons)", sec.axis = sec_axis(~./coeff, name = "Number of vessels landing market squid")) +
    theme(legend.position="bottom") + scale_x_continuous(name = element_blank()) +
    theme(plot.title = element_text(size=9, face="bold.italic"),
          axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
    guides(colour=guide_legend(title="Variables: ")) + scale_color_brewer(palette="Set2",
                     limits = c("LANDED_WEIGHT_MTONS.sum.sum", "n_vessel"),
                     labels=c("LANDED_WEIGHT_MTONS.sum.sum" = "Landings of Pacific sardine", 
                              "n_vessel" = "Number of vessels landing market squid")) 

# rm(g1, g2, g3, g4, df, landing.nvessels.year.sel, coeff)
# rm(landing.price.year, nvessel.year)

```


```{r ts_by_port,  fig.cap = "Total annual landing by selected ports during the period 2000 - 2020 for market squid and Pacific sardine. The ports selected are the ones with the highest average annual landings and substitution is likely.\\label{fig:ts_landings_by_ports}", echo=FALSE, message=FALSE, warning=FALSE}

landings.by.sel.ports <- landings.by.port.year %>% 
  filter(PACFIN_PORT_CODE == "SP" | PACFIN_PORT_CODE == "TRM" | 
         PACFIN_PORT_CODE == "MOS") %>%
  filter(PACFIN_SPECIES_CODE == "PSDN" |  PACFIN_SPECIES_CODE == "MSQD") %>%
  mutate(PACFIN_PORT_CODE = fct_relevel(PACFIN_PORT_CODE, "SP", "TRM", "HNM", "MNT", "MOS")) %>%
  filter(LANDING_YEAR >= 2000)

port_names <- as_labeller(c(`SP` = "San Pedro", `TRM` = "Terminal Island", 
                            `MOS` = "Moss Landing", `MNT` = "Monterrey", `HNM` = "Hueneme"))
     

ggplot(landings.by.sel.ports, aes(y=LANDED_WEIGHT_MTONS.sum.sum, 
                                  x=LANDING_YEAR, color=PACFIN_SPECIES_CODE, group=PACFIN_SPECIES_CODE)) +
  geom_line(size=1) + xlab("Year") + ylab("Landings (tons)") + 
  facet_wrap(~ PACFIN_PORT_CODE, labeller = port_names, nrow = 3, ncol = 1) +
  theme(legend.position="right") + 
  theme(axis.text.x = element_text(angle=90)) +
  guides(color=guide_legend(title="Species: ")) +
  scale_color_brewer(palette="Set2", labels=c("MSQD" = "Market Squid", "NANC" = "Northern Anchovy", 
             "PHRG" = "Pacific Herring", "PSDN" = "Pacific Sardine"))

rm(port_names, landings.by.sel.ports, landings.by.port.year)

```



# Participation (excluded from paper, enougth to use cluster)

```{r participation, eval=FALSE, include=FALSE}


total.revenue <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  group_by(VESSEL_NUM, LANDING_YEAR) %>% summarise(year_revenue = sum(AFI_EXVESSEL_REVENUE.sum))

### Vessel that participate in Pacific sardine fishery ###
psdn.fleet <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>%
  group_by(VESSEL_NUM, LANDING_YEAR) %>% summarise(revenue_psdn = sum(AFI_EXVESSEL_REVENUE.sum)) %>%
    merge(total.revenue, by = c("VESSEL_NUM", "LANDING_YEAR"), all.x = TRUE) %>%
  filter(revenue_psdn > 0) %>% group_by(VESSEL_NUM) %>%
  summarise(anual_revenue_psdn = mean(revenue_psdn), 
            anual_revenue_all = mean(year_revenue), 
            n_year_fishing = n()) %>%
  mutate(perc_psdn_rev = anual_revenue_psdn/anual_revenue_all) %>%
  filter(anual_revenue_psdn >= 1000) %>%
  filter(perc_psdn_rev >= 0.05) %>%
  filter(n_year_fishing >= 3) %>%
  select(VESSEL_NUM) %>% unique() %>% mutate(d_psdn_fleet = 1)

### Vessel that participate in Market squid fishery ###
msqd.fleet <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>%
  group_by(VESSEL_NUM, LANDING_YEAR) %>% summarise(revenue_msqd = sum(AFI_EXVESSEL_REVENUE.sum)) %>%
    merge(total.revenue, by = c("VESSEL_NUM", "LANDING_YEAR"), all.x = TRUE) %>%
  filter(revenue_msqd > 0) %>% group_by(VESSEL_NUM) %>%
  summarise(anual_revenue_msqd = mean(revenue_msqd), 
            anual_revenue_all = mean(year_revenue), 
            n_year_fishing = n()) %>%
  mutate(perc_msqd_rev = anual_revenue_msqd/anual_revenue_all) %>%
  filter(anual_revenue_msqd >= 1000) %>%
  filter(perc_msqd_rev >= 0.05) %>%
  filter(n_year_fishing >= 3) %>%
  select(VESSEL_NUM) %>% unique() %>% mutate(d_msqd_fleet = 1)

### Vessel that participate in Northen anchovy fishery ###
nanc.fleet <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("NANC")) %>%
  group_by(VESSEL_NUM, LANDING_YEAR) %>% summarise(revenue_nanc = sum(AFI_EXVESSEL_REVENUE.sum)) %>%
    merge(total.revenue, by = c("VESSEL_NUM", "LANDING_YEAR"), all.x = TRUE) %>%
  filter(revenue_nanc > 0) %>% group_by(VESSEL_NUM) %>%
  summarise(anual_revenue_nanc = mean(revenue_nanc), 
            anual_revenue_all = mean(year_revenue), 
            n_year_fishing = n()) %>%
  mutate(perc_nanc_rev = anual_revenue_nanc/anual_revenue_all) %>%
  filter(anual_revenue_nanc >= 1000) %>%
  filter(perc_nanc_rev >= 0.05) %>%
  filter(n_year_fishing >= 3) %>%
  select(VESSEL_NUM) %>% unique() %>% mutate(d_nanc_fleet = 1)

### Vessel that participate in Pacific mackerel fishery ###
cmck.fleet <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("CMCK")) %>%
  group_by(VESSEL_NUM, LANDING_YEAR) %>% summarise(revenue_cmck = sum(AFI_EXVESSEL_REVENUE.sum)) %>%
    merge(total.revenue, by = c("VESSEL_NUM", "LANDING_YEAR"), all.x = TRUE) %>%
  filter(revenue_cmck > 0) %>% group_by(VESSEL_NUM) %>%
  summarise(anual_revenue_cmck = mean(revenue_cmck), 
            anual_revenue_all = mean(year_revenue), 
            n_year_fishing = n()) %>%
  mutate(perc_cmck_rev = anual_revenue_cmck/anual_revenue_all) %>%
  filter(anual_revenue_cmck >= 1000) %>%
  filter(perc_cmck_rev >= 0.05) %>%
  filter(n_year_fishing >= 3) %>%
  select(VESSEL_NUM) %>% unique() %>% mutate(d_cmck_fleet = 1)

### Vessel that participate in Jack mackerel fishery ###
jmck.fleet <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("JMCK")) %>%
  group_by(VESSEL_NUM, LANDING_YEAR) %>% summarise(revenue_jmck = sum(AFI_EXVESSEL_REVENUE.sum)) %>%
    merge(total.revenue, by = c("VESSEL_NUM", "LANDING_YEAR"), all.x = TRUE) %>%
  filter(revenue_jmck > 0) %>% group_by(VESSEL_NUM) %>%
  summarise(anual_revenue_jmck = mean(revenue_jmck), 
            anual_revenue_all = mean(year_revenue), 
            n_year_fishing = n()) %>%
  mutate(perc_jmck_rev = anual_revenue_jmck/anual_revenue_all) %>%
  filter(anual_revenue_jmck >= 1000) %>%
  filter(perc_jmck_rev >= 0.05) %>%
  filter(n_year_fishing >= 3) %>%
  select(VESSEL_NUM) %>% unique() %>% mutate(d_jmck_fleet = 1)

### Vessel that participate in Unknown mackerel fishery ###
umck.fleet <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  filter(PACFIN_SPECIES_CODE  %in% c("UMCK")) %>%
  group_by(VESSEL_NUM, LANDING_YEAR) %>% summarise(revenue_umck = sum(AFI_EXVESSEL_REVENUE.sum)) %>%
    merge(total.revenue, by = c("VESSEL_NUM", "LANDING_YEAR"), all.x = TRUE) %>%
  filter(revenue_umck > 0) %>% group_by(VESSEL_NUM) %>%
  summarise(anual_revenue_umck = mean(revenue_umck), 
            anual_revenue_all = mean(year_revenue), 
            n_year_fishing = n()) %>%
  mutate(perc_umck_rev = anual_revenue_umck/anual_revenue_all) %>%
  filter(anual_revenue_umck >= 1000) %>%
  filter(perc_umck_rev >= 0.05) %>%
  filter(n_year_fishing >= 3) %>%
  select(VESSEL_NUM) %>% unique() %>% mutate(d_umck_fleet = 1)


# Select vessel that participate in the CPS fishery. 
PacFIN.month <- PacFIN.month %>% 
  merge(psdn.fleet, by = c("VESSEL_NUM"), all.y = FALSE, all.x = TRUE) %>%
  merge(msqd.fleet, by = c("VESSEL_NUM"), all.y = FALSE, all.x = TRUE) %>%
  merge(nanc.fleet, by = c("VESSEL_NUM"), all.y = FALSE, all.x = TRUE) %>%
  merge(cmck.fleet, by = c("VESSEL_NUM"), all.y = FALSE, all.x = TRUE) %>%
  merge(jmck.fleet, by = c("VESSEL_NUM"), all.y = FALSE, all.x = TRUE) %>% 
  filter(d_psdn_fleet == 1 | d_msqd_fleet == 1 | d_nanc_fleet == 1 | d_cmck_fleet == 1 | d_jmck_fleet == 1) %>%
  rowwise() %>% 
  mutate(d_fleet_CPS = sum(d_psdn_fleet, d_msqd_fleet, d_nanc_fleet, d_cmck_fleet, d_jmck_fleet, na.rm = TRUE))

rm(total.revenue, psdn.fleet, msqd.fleet, nanc.fleet, cmck.fleet, jmck.fleet, umck.fleet)

# CREATE NEW PACFIN.MONTH.CPS DATASET!
PacFIN.month.CPS <- PacFIN.month %>% 
    dplyr::filter(PACFIN_SPECIES_CODE %in% 
    c("CMCK", "JMCK", "MSQD", "NANC", "PSDN", "UMCK"))

```


# Clustering

```{r dataframe_cluster, include=FALSE}

Hierarchical_Vessel_Groups <- read.csv("C:\\GitHub\\EconAnalysis\\Clustering\\Hierarchical_Vessel_Groups.csv")
# Hierarchical_Vessel_Groups1 <- read.csv("C:\\GitHub\\EconAnalysis\\Clustering\\Hierarchical_Vessel_Groups1.csv")
# Hierarchical_Vessel_Groups2 <- read.csv("C:\\GitHub\\EconAnalysis\\Clustering\\Hierarchical_Vessel_Groups2.csv")
# Hierarchical_Vessel_Groups3 <- read.csv("C:\\GitHub\\EconAnalysis\\Clustering\\Hierarchical_Vessel_Groups3.csv")
# Hierarchical_Vessel_Groups4 <- read.csv("C:\\GitHub\\EconAnalysis\\Clustering\\Hierarchical_Vessel_Groups4.csv")

PacFIN.month.cluster <- merge(PacFIN.month, Hierarchical_Vessel_Groups, by = ("VESSEL_NUM"), all.x = TRUE, all.y = FALSE)
# PacFIN.month.cluster <- merge(PacFIN.month.cluster, Hierarchical_Vessel_Groups1, by = ("VESSEL_NUM"), all.x = TRUE, all.y = FALSE)
# PacFIN.month.cluster <- merge(PacFIN.month.cluster, Hierarchical_Vessel_Groups2, by = ("VESSEL_NUM"), all.x = TRUE, all.y = FALSE)
# PacFIN.month.cluster <- merge(PacFIN.month.cluster, Hierarchical_Vessel_Groups3, by = ("VESSEL_NUM"), all.x = TRUE, all.y = FALSE)
# PacFIN.month.cluster <- merge(PacFIN.month.cluster, Hierarchical_Vessel_Groups4, by = ("VESSEL_NUM"), all.x = TRUE, all.y = FALSE)

rm(Hierarchical_Vessel_Groups, Hierarchical_Vessel_Groups1, Hierarchical_Vessel_Groups2, 
   Hierarchical_Vessel_Groups3, Hierarchical_Vessel_Groups4,)

```

```{r cluster_descriptive}
library("googlesheets4")


gs4_auth(
    email = gargle::gargle_oauth_email(),
    path = NULL,
    scopes = "https://www.googleapis.com/auth/spreadsheets",
    cache = gargle::gargle_oauth_cache(),
    use_oob = gargle::gargle_oob_default(),
    token = NULL
)


### How gear are used by clusters??? ###
options(scipen=999)
cluster.gear <- PacFIN.month.cluster %>% filter(LANDING_YEAR >= 2000) %>% 
  group_by(group_all, PACFIN_GEAR_CODE) %>% summarise(landings = sum(LANDED_WEIGHT_MTONS.sum)) %>% 
  group_by(group_all) %>% mutate(Percentage = landings / sum(landings)) %>% 
  unique() %>% filter(group_all != is.na(group_all))

cluster.gear <- cluster.gear[order(cluster.gear$group_all, -cluster.gear$Percentage),]

# cluster.gear %>% 
#   group_by(group_all) %>% 
#   filter(row_number()==1:2)

cluster.gear <- cluster.gear %>%
  filter(PACFIN_GEAR_CODE %in% c("DPN", "ONT", "SEN", "CLP"))
    



table <- as.data.frame(xtabs(Percentage ~  PACFIN_GEAR_CODE + group_all, cluster.gear ))
table <- table %>%
  spread(key = group_all, value = Freq)

# table = table[,-1]
# rownames(table) = c("Crab and Lobster Pot", "Dip Net", "Other Net Gear", "Seine")
colnames(table) = c("Gear", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5", "Cluster 6", "Cluster 7")
gs4_create("Table1", sheets = table)
# print(xtable(table, caption = 'Percentage of cluster total langings by gear used.\\label{Table:cluster_gear}', type = "latex"), comment=FALSE,  caption.placement = "top")

rm(table, PacFIN.cluster)



#----------------------------------------------
##Step 7: Look at descriptive cluster metrics

###Look at the Port Groups Different Clusters are associated with based on % of total number of landings

Test1<-merge(Test2, Test3, by="group")
options(scipen=999) ## Avoid scientifi notation
Test1$Percentage<-Test1$AFI_EXVESSEL_REVENUE.x/Test1$AFI_EXVESSEL_REVENUE.y
Test1<-Test1[order(Test1$group, -Test1$Percentage),]

## Obtain two more important port areas by cluster
Test1<-
  Test1 %>% 
  group_by(group) %>% 
  filter(row_number()==1:2)

# Test1<-rbind(Test2, Test3)
Test1<-Test1[order(Test1$group, -Test1$Percentage),]
Test1

###Look at summary values for revenue and range
RAW$VESSEL_NUM<-Vessel_IDs
RAW<-merge(RAW, PAM_Vessel_Groups, by="VESSEL_NUM")
RAW<-RAW[c(-1)]
Group_Stats<-RAW%>% group_by(group) %>% summarise_each(funs(mean, se=sd(.)/sqrt(n())))
Group_Stats$DISTANCE_A_mean/10000
Group_Stats$AVG_REVENUE_mean

rm(Test1, Test2, Test3)


###Look at top Coastal Pelagic Species for Each Cluster
Test_Sub<-merge(Tickets, PAM_Vessel_Groups, by="VESSEL_NUM")
Test_Sub<-Test_Sub[which(Test_Sub$MGRP=="CPEL"),]

Test_Sub<- within(Test_Sub, Dominant[Dominant == "CHUB MACKEREL"] <- "OTHER")
Test_Sub<- within(Test_Sub, Dominant[Dominant == "JACK MACKEREL"] <- "OTHER")
Test_Sub<- within(Test_Sub, Dominant[Dominant == "UNSP. MACKEREL"] <- "OTHER")
Test_Sub<- within(Test_Sub, Dominant[Dominant == "ROUND HERRING"] <- "OTHER")
Test_Sub<- within(Test_Sub, Dominant[Dominant == "PACIFIC BONITO"] <- "OTHER")

Test1<-aggregate(AFI_EXVESSEL_REVENUE~Dominant+group, data=Test_Sub, FUN=sum)
Test2<-aggregate(AFI_EXVESSEL_REVENUE~group, data=Test_Sub, FUN=sum)
Test1<-merge(Test1, Test2, by="group")
options(scipen=999)
Test1$Percentage<-Test1$AFI_EXVESSEL_REVENUE.x/Test1$AFI_EXVESSEL_REVENUE.y
Test1<-Test1[order(Test1$group, -Test1$Percentage),]
Test1<-
  Test1 %>% 
  group_by(group) %>% 
  filter(row_number()==1:2)

Test1<-Test1[order(Test1$group, -Test1$Percentage),]
Test1

###Look at top non-Coastal Pelagic Species for Each Cluster
Test_Sub<-merge(Tickets, PAM_Vessel_Groups, by="VESSEL_NUM")
Test_Sub<-Test_Sub[which(!Test_Sub$MGRP=="CPEL"),]
Test1<-aggregate(AFI_EXVESSEL_REVENUE~MGRP+group, data=Test_Sub, FUN=sum)
Test2<-aggregate(AFI_EXVESSEL_REVENUE~group, data=Test_Sub, FUN=sum)
Test1<-merge(Test1, Test2, by="group")
options(scipen=999)
Test1$Percentage<-Test1$AFI_EXVESSEL_REVENUE.x/Test1$AFI_EXVESSEL_REVENUE.y
Test1<-Test1[order(Test1$group, -Test1$Percentage),]
Test1<-
  Test1 %>% 
  group_by(group) %>% 
  filter(row_number()==1:2)

Test1<-Test1[order(Test1$group, -Test1$Percentage),]
Test1

# ###Look at number of Vessels Having CPS Permits
# CPS_Permits<-read.csv("CPS_Permits.csv")
# CPS_Permits<-merge(CPS_Permits, PAM_Vessel_Groups, by="VESSEL_NUM", all.x=TRUE)
# CPS_Permits<-aggregate(VESSEL_NUM~group, FUN=length, data=CPS_Permits)
# Group_Totals<-aggregate(VESSEL_NUM~group, FUN=length, data=PAM_Vessel_Groups)
# CPS_Permits<-merge(CPS_Permits, Group_Totals, by="group")
# CPS_Permits$Percentage<-CPS_Permits$VESSEL_NUM.x/CPS_Permits$VESSEL_NUM.y
# CPS_Permits



```
<!-- CHECK FROM HERE!!! -->



# Switching behavior

Substitution between species is likely related to the gear that a vessel or fisher have invested in for their fishing activities and the port they commonly land. These two characteristics will dictate the potential species choices that a vessel has to choose. @richerson2017 call species choices and other spatial patterns of ocean use as spatial constraints. Vessels with the same spatial constraints share local ecological knowledge and vessel mobility [@richerson2017]. Nevertheless, vessels are not restricted to a specific gear and port. In theory, they can switch between different gear alternatives and choose different ports to land. Therefore, they can modify the spatial constraints that they face. Nevertheless, switching between gears that are entirely different could be costly, and landing in a port without the infrastructure to land the targeted is unlikely.

In the case of CPS species, vessels that use nets as gear for harvest only require to change net sizes, which is an easy task. Therefore, we should observe higher substitution within these vessels that uses net gears. We can even observe landing on the same day of Pacific sardine and market squid as the former is a day fishery, while the latter is a night one. Regrading to ports, substitution between Pacific sardine and market squid is more likely in the south of the California Current System (CCS) as both species have a fishery running around this area. Specifically, substitution between market squid and Pacific sardine seems more likely in Los Angeles, Monterrey, and Santa Barbara area ports as positive values for market squid and Pacific sardine are observed. Meanwhile, there is less target switching in the north of the CCS (CPS Workshop).  

Vessel traveling considerable distances to catch CPS species is unlikely. Because CPS spoils quickly, fishers stay close to ports and usually come back to port on the same day. Therefore, ports constraints could have consequences if species distribution shifts north due to climate conditions. Probably, fishers will not target market squid, independently of its biomass levels, if fishing grounds move north,  as there is no close infrastructure for its landing (CPS workshop). Nevertheless, there is uncertainty about this assumption as it would depend on the speed that ports can anticipate and adjust their landings capacity and infrastructure in the future. In this section, we analyze our dataset to understand better how gears and ports are related to switching between species and whether individual vessels switch gears and ports during the historical data.


### Switching by gear used

```{r n_species, eval=FALSE, include=FALSE}
PacFIN.species <- PacFIN.month %>% select('PACFIN_SPECIES_CODE') %>% unique()
nrow(PacFIN.species)

```

In our sample data, a vessel belonging to the CPS fleet uses 2.194 gears on average to harvest CPS species, with a standard deviation of 1.01. <!-- If we disaggregate this information by species harvested by vessels that participate in the corresponding fishery, a vessel switches between 1.87 gears on average to harvest market squid, 1.50 gears on average to harvest Northern anchovy, 1.80 gears on average to harvest Pacific mackerel, and 1.56 on average to harvest Pacific sardine. In terms of gears,  --> Seine and other net gears (e.g., round-haul nets) allow for the highest substitution level between all CPS species harvested by the CPS fleet. A vessel using seine harvest has an average of 3.55 CPS species, while a vessel using other net gear harvests 2.95 CPS species. Regarding to species, seine is the most used gear to harvest Pacific sardine, Northern anchovy, market squid, and Pacific mackerel by their corresponding fleet. 60\% of vessels have used seine to harvest Pacific mackerel since 2000,  94\% to harvest market squid, 90\% to harvest Pacific sardine, and 94\% to harvest Northern anchovy. <!-- Regarding total landings, seine is the most important gear by far. The percentage of the total landing that comes from using seine is 97\% for market squid, 96\%  for Northern anchovy and 93\%  for Pacific mackerel, and 67\% for the Pacific sardine. The other 33\% of the total landings of Pacific sardine come from other net gears -->

```{r n_gears_used_vessels, eval=FALSE, include=FALSE}
# How many gears a vessel use in the dataset
PacFIN.CPS.unique.gear <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_GEAR_CODE) %>% unique()

PacFIN.CPS.n_gears <- PacFIN.CPS.unique.gear %>% mutate(n_gears = 1) %>% group_by(VESSEL_NUM) %>% 
  summarise(n_gears= sum(n_gears)) 
  summary(PacFIN.CPS.n_gears$n_gears)
  sd(PacFIN.CPS.n_gears$n_gears)

# ### If I have used SEINE, how many other gear I use?
# PacFIN.Base <- PacFIN_month %>% filter( LANDING_YEAR >= 2010) %>%  select(VESSEL_NUM, PACFIN_GEAR_CODE) %>% unique()
#   PacFIN.only.seine <- PacFIN_month %>% filter( LANDING_YEAR >= 2010) %>% select(VESSEL_NUM, PACFIN_GEAR_CODE) %>%
#     filter(PACFIN_GEAR_CODE == "SEN") %>% mutate(D_Seine = 1) %>% select(VESSEL_NUM, D_Seine) %>% unique() 
#   PacFIN.CPS.n_gear <- PacFIN.Base %>% inner_join(PacFIN.only.seine, by = c("VESSEL_NUM")) %>% filter(D_Seine == 1) %>% 
#     unique() %>% group_by(VESSEL_NUM) %>% summarise(N_gear= sum(D_Seine)) 
#   summary(PacFIN.CPS.n_gear$N_gear)
#   sd(PacFIN.CPS.n_gear$N_gear)
  
rm(PacFIN.CPS.unique.gear, PacFIN.CPS.n_gears)

```

```{r n_gear_by_species, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
PacFIN.Gear.PSDN <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, d_psdn_fleet) %>% unique() %>% 
  filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>% filter(d_psdn_fleet == 1) %>% mutate(n_gears = 1) %>%  
  group_by(VESSEL_NUM) %>% summarise(n_gears = sum(n_gears)) %>% mutate(Species = "PSDN")

PacFIN.Gear.MSQD <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, d_msqd_fleet) %>% unique() %>% 
  filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>% filter(d_msqd_fleet == 1) %>% mutate(n_gears = 1) %>%  
  group_by(VESSEL_NUM) %>% summarise(n_gears = sum(n_gears)) %>% mutate(Species = "MSQD")

PacFIN.Gear.NANC <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, d_nanc_fleet) %>% unique() %>% 
  filter(PACFIN_SPECIES_CODE  %in% c("NANC")) %>% filter(d_nanc_fleet == 1) %>% mutate(n_gears = 1) %>%  
  group_by(VESSEL_NUM) %>% summarise(n_gears = sum(n_gears)) %>% mutate(Species = "NANC")

PacFIN.Gear.CMCK <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, d_cmck_fleet) %>% unique() %>% 
  filter(PACFIN_SPECIES_CODE  %in% c("CMCK")) %>% filter(d_cmck_fleet == 1) %>% mutate(n_gears = 1) %>%  
  group_by(VESSEL_NUM) %>% summarise(n_gears = sum(n_gears)) %>% mutate(Species = "CMCK")

PacFIN.Gear <- rbind.data.frame(PacFIN.Gear.PSDN, PacFIN.Gear.NANC, 
                                PacFIN.Gear.MSQD, PacFIN.Gear.CMCK)

table <- xtabs(n_gears ~ Species, aggregate(n_gears ~ Species, PacFIN.Gear, mean))
rownames(table) = c("Pacific Mackerel", "Market squid", "Northern Anchovy", "Pacific sardine") 
print(xtable(table, caption = 'Average number of gear used to catch a CPS species.', type = "latex"), comment=FALSE,  caption.placement = "top")

rm(PacFIN.Gear, PacFIN.Gear.PSDN, PacFIN.Gear.NANC, PacFIN.Gear.MSQD, PacFIN.Gear.PHRG, PacFIN.Gear.CMCK)

```

```{r n_species_species_gear, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}

PacFIN.CPS <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE) %>% 
  unique() %>%  
  mutate(CPS_species = 1) %>%
  group_by(VESSEL_NUM, PACFIN_GEAR_CODE) %>% 
  summarise(n_CPS_landed = sum(CPS_species)) %>%
  filter(PACFIN_GEAR_CODE %in% c("DPN", "ONT", "SEN", "TRL", "POL")) 

table <- xtabs(n_CPS_landed ~ PACFIN_GEAR_CODE, aggregate(n_CPS_landed ~ PACFIN_GEAR_CODE, PacFIN.CPS, mean))
table

rownames(table) = c("Dip Net", "Other Net Gear", "Pole", "Seine", "Troll") 

print(xtable(table, caption = 'Average number of CPS species landed by a vessel by gear used.\\label{Table:n_CPS_gear}', type = "latex"), comment=FALSE,  caption.placement = "top")

rm(PacFIN.CPS, table)

```

```{r n_vessels_using_gear, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
PacFIN.Gear.PSDN <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_psdn_fleet == 1) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(n_vessels = n()) %>% mutate(Species = "PSDN") %>% 
  filter(PACFIN_GEAR_CODE %in% c("DPN", "ONT", "POL", "SEN", "TRL"))
  n_psdn <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_psdn_fleet == 1) %>% 
    select(VESSEL_NUM, PACFIN_SPECIES_CODE) %>% 
    unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>% select(VESSEL_NUM) %>% unique()

PacFIN.Gear.MSQD <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_msqd_fleet == 1) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(n_vessels = n()) %>% mutate(Species = "MSQD") %>% 
  filter(PACFIN_GEAR_CODE %in% c("DPN", "ONT", "POL", "SEN", "TRL"))
  n_msqd <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_msqd_fleet == 1) %>%
    select(VESSEL_NUM, PACFIN_SPECIES_CODE) %>% 
    unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>% select(VESSEL_NUM) %>% unique()
  
PacFIN.Gear.NANC <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_nanc_fleet == 1) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE %in% c("NANC")) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(n_vessels = n()) %>% mutate(Species = "NANC") %>% 
  filter(PACFIN_GEAR_CODE %in% c("DPN", "ONT", "POL", "SEN", "TRL"))
  n_nanc <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_nanc_fleet == 1) %>%
    select(VESSEL_NUM, PACFIN_SPECIES_CODE) %>% 
    unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("NANC")) %>% select(VESSEL_NUM) %>% unique()
  
PacFIN.Gear.CMCK <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_cmck_fleet == 1) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE %in% c("CMCK")) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(n_vessels = n()) %>% mutate(Species = "CMCK") %>% 
  filter(PACFIN_GEAR_CODE %in% c("DPN", "ONT", "POL", "SEN", "TRL"))
  n_cmck <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_cmck_fleet == 1) %>%
    select(VESSEL_NUM, PACFIN_SPECIES_CODE) %>% 
    unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("CMCK")) %>% select(VESSEL_NUM) %>% unique()
  
PacFIN.Gear.PSDN$n_vessels <- PacFIN.Gear.PSDN$n_vessels/nrow(n_psdn)
PacFIN.Gear.MSQD$n_vessels <- PacFIN.Gear.MSQD$n_vessels/nrow(n_msqd)
PacFIN.Gear.NANC$n_vessels <- PacFIN.Gear.NANC$n_vessels/nrow(n_nanc)
PacFIN.Gear.CMCK$n_vessels <- PacFIN.Gear.CMCK$n_vessels/nrow(n_cmck)
  PacFIN.Gear <- rbind.data.frame(PacFIN.Gear.PSDN, PacFIN.Gear.NANC, PacFIN.Gear.MSQD, PacFIN.Gear.CMCK)
  table <- xtabs(n_vessels ~ PACFIN_GEAR_CODE + Species, PacFIN.Gear)

colnames(table) = c("Pacific mackerel", "Market squid", "Northern anchovy", "Pacific sardine") 
  rownames(table) = c("Dip Net", "Other Net Gear", "Pole", "Seine", "Troll") 

print(xtable(table, 
             caption = 'Percentage of vessels using a gear to harvest CPS species.\\label{table:perc_n_vessel}', 
             type = "latex"), comment=FALSE,  caption.placement = "top")

rm(PacFIN.Gear, PacFIN.Gear.PSDN, PacFIN.Gear.NANC, PacFIN.Gear.MSQD, PacFIN.Gear.CMCK,
   table, n_cmck, n_msqd, n_nanc, n_psdn)

```

```{r landing_using_gear, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
PacFIN.Gear.PSDN <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, LANDED_WEIGHT_MTONS.sum) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE %in% c("PSDN")) %>% 
  mutate(tot_landings = sum(LANDED_WEIGHT_MTONS.sum)) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(landings = sum(LANDED_WEIGHT_MTONS.sum), tot_landings = mean(tot_landings)) %>% 
  mutate(Species = "PSDN") %>% filter(PACFIN_GEAR_CODE %in% c("DPN", "MDT", "ONT", "SEN"))

PacFIN.Gear.MSQD <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, LANDED_WEIGHT_MTONS.sum) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE %in% c("MSQD")) %>% 
  mutate(tot_landings = sum(LANDED_WEIGHT_MTONS.sum)) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(landings = sum(LANDED_WEIGHT_MTONS.sum), tot_landings = mean(tot_landings)) %>% 
  mutate(Species = "MSQD") %>% filter(PACFIN_GEAR_CODE %in% c("DPN", "MDT", "ONT", "SEN"))

PacFIN.Gear.NANC <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, LANDED_WEIGHT_MTONS.sum) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE %in% c("NANC")) %>% 
  mutate(tot_landings = sum(LANDED_WEIGHT_MTONS.sum)) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(landings = sum(LANDED_WEIGHT_MTONS.sum), tot_landings = mean(tot_landings)) %>% 
  mutate(Species = "NANC") %>% filter(PACFIN_GEAR_CODE %in% c("DPN", "MDT", "ONT", "SEN"))

PacFIN.Gear.CMCK <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% 
  select(PACFIN_SPECIES_CODE, PACFIN_GEAR_CODE, LANDED_WEIGHT_MTONS.sum) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE %in% c("CMCK")) %>% 
  mutate(tot_landings = sum(LANDED_WEIGHT_MTONS.sum)) %>% group_by(PACFIN_GEAR_CODE) %>%
  summarise(landings = sum(LANDED_WEIGHT_MTONS.sum), tot_landings = mean(tot_landings)) %>% 
  mutate(Species = "CMCK") %>% filter(PACFIN_GEAR_CODE %in% c("DPN", "MDT", "ONT", "SEN"))

PacFIN.Gear.PSDN$n_vessels <- PacFIN.Gear.PSDN$landings/PacFIN.Gear.PSDN$tot_landings
PacFIN.Gear.MSQD$n_vessels <- PacFIN.Gear.MSQD$landings/PacFIN.Gear.MSQD$tot_landings
PacFIN.Gear.NANC$n_vessels <- PacFIN.Gear.NANC$landings/PacFIN.Gear.NANC$tot_landings
PacFIN.Gear.CMCK$n_vessels <- PacFIN.Gear.CMCK$landings/PacFIN.Gear.CMCK$tot_landings
  PacFIN.Gear <- rbind.data.frame(PacFIN.Gear.PSDN, PacFIN.Gear.NANC, PacFIN.Gear.MSQD, PacFIN.Gear.CMCK)
  table <- xtabs(n_vessels ~ PACFIN_GEAR_CODE + Species, PacFIN.Gear)

colnames(table) = c("Pacific mackerel", "Market squid", "Northern anchovy", "Pacific sardine") 
  rownames(table) = c("Dip Net", "Midwater Trawl", "Other Net Gear", "Seine") 

print(xtable(table, 
    caption = 'Percentage of landing by gear used and CPS species harvested.\\label{table:perc_landings}', 
    type = "latex"), comment=FALSE,  caption.placement = "top")

rm(PacFIN.Gear, PacFIN.Gear.PSDN, PacFIN.Gear.NANC, PacFIN.Gear.MSQD, table)

```

The ability to switch between CPS species using seine as gear is confirmed when we analyze the most important gears used by different clusters in terms of landings (see Table \ref{Table:cluster_gear}). Total landings of a cluster where substitution between CPS species is observable (Cluster 2 to 5) comes mainly from catches using seine, and in some cases, from other net gears. In the case of Cluster 6 and 7, where substitution is between a CPS species and a Non-CPS species, seine or Other Net Gear are not important, and other gears like Dip Net or Troll became relevant.

### Switching by port landed

```{r n_ports_vessels_CPS, eval=FALSE, include=FALSE}
# How many gears a vessel use in the dataset
PacFIN.CPS.unique.port <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_PORT_CODE) %>% unique() 
  PacFIN.CPS.n_ports <- PacFIN.CPS.unique.port %>% mutate(n_ports = 1) %>% 
    group_by(VESSEL_NUM) %>% summarise(n_ports= sum(n_ports)) 
    summary(PacFIN.CPS.n_ports$n_ports)
    sd(PacFIN.CPS.n_ports$n_ports)

rm(PacFIN.CPS.unique.port, PacFIN.CPS.n_ports)

```

```{r n_ports_CPS, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
PacFIN.Gear.PSDN <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_psdn_fleet == 1) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_PORT_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>% 
  mutate(N_port = 1) %>% group_by(VESSEL_NUM) %>% 
  summarise(N_port = sum(N_port)) %>% mutate(Species = "PSDN")

PacFIN.Gear.MSQD <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_msqd_fleet == 1) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_PORT_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>% 
  mutate(N_port = 1) %>% group_by(VESSEL_NUM) %>% 
  summarise(N_port = sum(N_port)) %>% mutate(Species = "MSQD")

PacFIN.Gear.NANC <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_nanc_fleet == 1) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_PORT_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("NANC")) %>% 
  mutate(N_port = 1) %>% group_by(VESSEL_NUM) %>% 
  summarise(N_port = sum(N_port)) %>% mutate(Species = "NANC")

PacFIN.Gear.CMCK <- PacFIN.month %>% filter(LANDING_YEAR >= 2000) %>% filter(d_cmck_fleet == 1) %>% 
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, PACFIN_PORT_CODE) %>% 
  unique() %>% filter(PACFIN_SPECIES_CODE  %in% c("CMCK")) %>% 
  mutate(N_port = 1) %>% group_by(VESSEL_NUM) %>% 
  summarise(N_port = sum(N_port)) %>% mutate(Species = "CMCK")

PacFIN.Port <- rbind.data.frame(PacFIN.Gear.PSDN, PacFIN.Gear.NANC,
                                PacFIN.Gear.MSQD, PacFIN.Gear.CMCK)

table <- xtabs(N_port ~ Species, aggregate(N_port ~ Species, PacFIN.Port, mean))
rownames(table) = c("Pacific mackerel", "Market squid", "Northern Anchovy", "Pacific sardine") 
print(xtable(table, caption = 'Average number of ports where a vessel landed their CPS species catch.', 
             type = "latex"), comment=FALSE,  caption.placement = "top")

rm(PacFIN.Gear.PSDN, PacFIN.Gear.MSQD, PacFIN.Gear.NANC, PacFIN.Gear.CMCK,  PacFIN.Port)

```

On average, vessels participating in the CPS fishery land their CPS catch in 5.36 ports, with a standard deviation of 3.65. When we disaggregate this information by species, vessels that participate in the corresponding fishery land on average market squid in 5.37 ports, Pacific sardine in 3.96 ports, Northern anchovy in 3.17 ports, and Pacific mackerel in 2.24 ports. 

Using the results from the cluster analysis, we can identify how many ports, on average, a vessel that belongs to a cluster lands their CPS catch. On average, vessels belonging to Cluster 1 land CPS in 4.81 ports, Cluster 2 land CPS  in 5.29 ports, Cluster 3 land CPS in 5.06 ports, Cluster 4 land CPS in 7.71 ports, Cluster 5 land in 3.5 ports, Cluster 6 land in CPS 2.88, and Cluster 7 land in CPS  2.22. Finally, Table \ref{Table:cluster_port} show the percentage of a cluster´s total CPS landing that is landed in a port.

```{r n_ports_CPS_cluster, eval=FALSE, include=FALSE}
# How many gears a vessel use in the dataset
PacFIN.CPS.unique.port <- PacFIN.month.CPS.cluster %>% filter(LANDING_YEAR >= 2000) %>% 
  select(VESSEL_NUM, PACFIN_PORT_CODE, cluster) %>% unique() 
  PacFIN.CPS.n_ports <- PacFIN.CPS.unique.port %>% mutate(n_ports = 1) %>% 
    group_by(VESSEL_NUM, cluster) %>% summarise(n_ports= sum(n_ports)) %>% 
    group_by(cluster) %>% summarise(n_ports = mean(n_ports))
  
rm(PacFIN.CPS.unique.port, PacFIN.CPS.n_ports)

```

```{r port_by_cluster, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Create dummy of whether a vessel harvest MSQD, NANC OR PSDN 
PacFIN.cluster <- PacFIN.month.CPS.cluster %>% filter(LANDING_YEAR >= 2000) %>% 
  group_by(group, PACFIN_PORT_CODE) %>% summarise(landings = sum(LANDED_WEIGHT_MTONS.sum)) %>% 
  group_by(group) %>% mutate(perc = landings / sum(landings)) %>% 
  unique() %>% filter(group != is.na(group)) 

# %>% 
#   filter(PACFIN_PORT_CODE %in% 
#       c("SP", "TRM", "HNM", "VEN", "MNT", "MOS", "AST", "LWC", "WPT")) %>%
#   mutate(PACFIN_PORT_CODE = fct_relevel(PACFIN_PORT_CODE, 
#       "SP", "TRM", "HNM", "VEN", "MNT", "MOS", "AST", "LWC", "WPT"))
    
table <- xtabs(perc ~  PACFIN_PORT_CODE + group, PacFIN.cluster)
# rownames(table) = c("San Pedro", "Terminal Island", "Hueneme", "Ventura", "Monterrey", "Moss Landing", 
#                     "Astoria", "Ilwaco / Chinook", "Westport")
  
colnames(table) = c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")

print(xtable(table, caption = 'Percentage of cluster total landings by port.\\label{Table:cluster_port}', type = "latex"), comment=FALSE,  caption.placement = "top")

rm(table, PacFIN.cluster)

```

<!-- ### Vessel lenght  -->
<!-- <!-- 1) What are the characteristics of a vessel depending on gear, and also on Species -->
<!-- There is heterogeneity on vessel characteristics depending on the gear and target species. In general, vessel harvesting Pacific herring tends to be smaller in length (see Table \ref{Table:length}), except for vessels using pole. Vessels using seine and harvesting Market squid, Pacific sardine and Northern anchovy have similar length, suggesting easier substitution between these species. About fishing days... -->

<!-- ```{r lenght_by_species_gear, echo=FALSE, results='asis'} -->

<!-- PacFIN.CPS <- PacFIN.month %>%  -->
<!--   filter(PACFIN_SPECIES_CODE  %in% c("MSQD", "PSDN", "NANC", "PHRG"))  %>%  -->
<!--   filter(PACFIN_GEAR_CODE  %in% c("DPN", "ONT", "SEN", "GLN", "MDT", "POL")) %>% -->
<!--   filter(LANDING_YEAR >= 2010 ) -->

<!-- table <- xtabs(VESSEL_LENGTH.mean ~ PACFIN_GEAR_CODE + PACFIN_SPECIES_CODE, -->
<!--                aggregate(VESSEL_LENGTH.mean ~ PACFIN_GEAR_CODE + PACFIN_SPECIES_CODE, PacFIN.CPS, mean)) -->
<!-- # table -->
<!-- colnames(table) = c("Market squid", "Northern Anchovy", "Pacific Herring", "Pacific Sardine")  -->
<!-- rownames(table) = c("Dip Net", "Gill Net", "Midwater Trawl", "Other Net Gear", "Pole", "Seine")  -->
<!-- print(xtable(table, caption = 'Vessel lenght used in the CPS fishery by gear used.\\label{Table:lenght}', type = "latex"), comment=FALSE,  caption.placement = "top") -->

<!-- rm(table, PacFIN.CPS) -->

<!-- ``` -->


<!-- ### Fishing days -->

<!-- ```{r fishing_days, echo=FALSE, results='asis'} -->

<!-- PacFIN.CPS <- PacFIN.month %>%  -->
<!--   filter(PACFIN_SPECIES_CODE  %in% c("MSQD", "PSDN", "NANC", "PHRG"))  %>%  -->
<!--   filter(PACFIN_GEAR_CODE  %in% c("DPN", "ONT", "SEN", "GLN", "MDT", "POL")) %>% -->
<!--   filter(LANDING_YEAR >= 2000 ) -->

<!-- table <- xtabs(NUM_OF_DAYS_FISHED.mean ~ PACFIN_GEAR_CODE + PACFIN_SPECIES_CODE, -->
<!--                aggregate(NUM_OF_DAYS_FISHED.mean ~ PACFIN_GEAR_CODE + PACFIN_SPECIES_CODE, PacFIN.CPS, mean)) -->
<!--  # table -->
<!-- colnames(table) = c("Northern Anchovy", "Pacific Herring", "Pacific Sardine")  -->
<!-- rownames(table) = c("Dip Net", "Midwater Trawl", "Other Net Gear", "Seine") -->
<!-- print(xtable(table, caption = 'Number of days fishing by species and gear used', type = "latex"), comment=FALSE,  caption.placement = "top") -->

<!-- rm(table, PacFIN.CPS) -->

<!-- ``` -->


### Other determinants switching of behavior

Switching behavior might depend on whether a vessel is owner-operated or company-owned (CPS Workshop). Owner-operated vessels can freely choose what to harvest (and when and where), while company-owned vessels are restricted by what the company dictates. However, regardless of vessel ownership, most fishing behavior would be determined by market order (CPS workshop). Vessels will already have contracts with processors before fishing (CPS workshop), deciding on the target species and landing port ex-ante. Nevertheless, if favorable conditions arise, the skipper has some freedom to switch species. Therefore, we expect more flexibility in species substitution but more rigidity on port landings unless a processor owns different landing facilities, allowing vessels to increase their flexibility regarding port landing. 

Within the restrictions imposed by contracts, revenues and harvest costs will affect the target decision. Price, in general, varies by port of landing and harvest cost by the distance traveled to fishing grounds. It is an empirical question whether prices and distances directly affect the decision to target a species, or fishers consider the relative prices and distance to decide with species to harvest.  

<!-- Another determinant of behavior that vessels face is the perishability of the catch. CPS vessels are limited to staying at sea for one day because CPS species can perish if stored for longer. It does not mean that vessels cannot target different species in one day. Some vessels harvest market squid and Pacific sardine during the same day.  -->

```{r switching_day_landings, eval=FALSE, include=FALSE}
# PacFIN_2010_2020 <- read.csv(file = "C:/Data/PacFIN data/FutureSeasIII_2010_2020.csv")
#   PacFIN_2000_2009 <- read.csv(file = "C:/Data/PacFIN data/FutureSeasIII_2000_2009.csv")
#
#   # Merge different decades of data #
# PacFIN <- rbind.data.frame(PacFIN_2000_2009, PacFIN_2010_2020)
#   rm(PacFIN_2000_2009, PacFIN_2010_2020)
#   gc()
#
# psdn.landings.by.day.vessel <- PacFIN %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
#   select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR, LANDING_MONTH, LANDING_DAY, LANDED_WEIGHT_MTONS) %>%
#   filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>%
#   group_by(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR, LANDING_MONTH, LANDING_DAY) %>%
#   summarise(Landed_PSDN = sum(LANDED_WEIGHT_MTONS)) %>% mutate(harv.psdn = 1)
#
# # If a vessel harvest Pacific sardine in a year, how many also harvest squid.
# squid.cond.psdn <- PacFIN %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
#   select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR, LANDING_MONTH, LANDING_DAY, LANDED_WEIGHT_MTONS) %>%
#   merge(psdn.landings.by.day.vessel, by = c("VESSEL_NUM", "LANDING_YEAR", "LANDING_MONTH", "LANDING_DAY"), all.x = TRUE) %>%
#   filter(harv.psdn == 1) %>% filter(PACFIN_SPECIES_CODE.x == "MSQD") %>%
#   mutate(perc_land_squid = LANDED_WEIGHT_MTONS / (LANDED_WEIGHT_MTONS + Landed_PSDN))
#
# write.csv(squid.cond.psdn,"C:\\Data\\day_switch_land.csv", row.names = FALSE)
df <- read.csv(file = "C:\\Data\\day_switch_land.csv")
hist(df$perc_land_squid, col = 'skyblue3', breaks = 10)
plot.ecdf(df$perc_land_squid)

```

```{r switching_day, eval=FALSE, include=FALSE}
# PacFIN_2010_2020 <- read.csv(file = "C:/Data/PacFIN data/FutureSeasIII_2010_2020.csv")
#   PacFIN_2000_2009 <- read.csv(file = "C:/Data/PacFIN data/FutureSeasIII_2000_2009.csv")
#
#   # Merge different decades of data #
# PacFIN <- rbind.data.frame(PacFIN_2000_2009, PacFIN_2010_2020)
#   rm(PacFIN_2000_2009, PacFIN_2010_2020)
#   gc()
#
# psdn.by.day <- PacFIN %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
#   select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR, LANDING_MONTH, LANDING_DAY) %>%
#   filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>%
#   unique() %>% mutate(harv.psdn = 1)
#
# # If a vessel harvest Pacific sardine in a year, how many also harvest squid.
# squid.cond.psdn <- PacFIN %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
#   select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR, LANDING_MONTH, LANDING_DAY) %>%
#   merge(psdn.by.day, by = c("VESSEL_NUM", "LANDING_YEAR", "LANDING_MONTH", "LANDING_DAY"), all.x = TRUE) %>%
#   filter(harv.psdn == 1) %>% filter(PACFIN_SPECIES_CODE.x == "MSQD") %>% unique()
#
# # Calculate number of vessels harvesting sardine and squid and percentage per year
# n_vessel_psdn <- psdn.by.day %>% group_by(LANDING_YEAR, LANDING_MONTH, LANDING_DAY) %>%
#   summarize(n_vessel_psdn = sum(harv.psdn, na.rm = TRUE))
# # %>% group_by(LANDING_MONTH, LANDING_DAY) %>%
# #   summarize(n_vessel_psdn = mean(n_vessel_psdn, na.rm = TRUE))
#
# n_vessel_msqd_c_psdn <- squid.cond.psdn %>% group_by(LANDING_YEAR, LANDING_MONTH, LANDING_DAY) %>%
#   summarize(n_vessel_msqd = sum(harv.psdn, na.rm = TRUE))
# # %>% group_by(LANDING_MONTH, LANDING_DAY) %>%
# #   summarize(n_vessel_msqd = mean(n_vessel_msqd, na.rm = TRUE))
#
# df <- merge(n_vessel_psdn, n_vessel_msqd_c_psdn, by=c("LANDING_YEAR", "LANDING_MONTH", "LANDING_DAY"), all.x = TRUE) %>%
#     mutate(n_vessel_msqd = if_else(is.na(n_vessel_msqd), 0, n_vessel_msqd)) %>% group_by(LANDING_MONTH, LANDING_DAY) %>%
#     summarize(n_vessel_msqd = sum(n_vessel_msqd, na.rm = TRUE), n_vessel_psdn = sum(n_vessel_psdn, na.rm = TRUE)) %>%
#     mutate(perc = (n_vessel_msqd / n_vessel_psdn)) %>% select(LANDING_MONTH, LANDING_DAY, perc)
#
# df$MonthDay <- paste( month.abb[df$LANDING_MONTH], df$LANDING_DAY, sep="-" )
#
# write.csv(df,"C:\\Data\\day_switch.csv", row.names = FALSE)
df <- read.csv(file = "C:\\Data\\day_switch.csv")
ggplot(data=df, aes(x=MonthDay, y=perc)) +
    geom_bar(stat="identity", fill=rgb(0.1,0.4,0.5,0.7), width=0.4) +
  scale_x_discrete(name = "Day", breaks=1:366) +
  scale_y_percent(name = "Percentage of vessels harvesting squid" )
rm(df)
```

```{r switching_graph, eval=FALSE, fig.cap="Switching", message=FALSE, warning=FALSE, include=FALSE}
psdn.by.year <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR) %>% filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>%
  unique() %>% mutate(harv.psdn = 1)
msqd.by.year <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR) %>% filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>%
  unique() %>% mutate(harv.msqd = 1)

# If a vessel harvest Pacific sardine in a year, how many also harvest squid.
squid.cond.psdn <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR) %>% merge(psdn.by.year,
                        by = c("VESSEL_NUM", "LANDING_YEAR"), all.y = TRUE) %>%
  filter(harv.psdn == 1) %>% filter(PACFIN_SPECIES_CODE.x == "MSQD") %>% unique()

# Calculate number of vessels harvesting sardine and squid and percentage per year
n_vessel_psdn <- psdn.by.year %>% group_by(LANDING_YEAR) %>% summarize(n_vessel_psdn = sum(harv.psdn))
n_vessel_msqd_c_psdn <- squid.cond.psdn %>% group_by(LANDING_YEAR) %>% summarize(n_vessel_msqd = sum(harv.psdn))
  df <- n_vessel_msqd_c_psdn %>% mutate(perc = n_vessel_psdn$n_vessel_psdn) %>%
    mutate(n_vessel_only_psdn = perc) %>%  # - n_vessel_msqd
    mutate(perc = (n_vessel_msqd / perc))

coeff = 115
g1 <- ggplot(df) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = n_vessel_only_psdn,
                          color = "# of vessels landing sardine"), size = 1) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = n_vessel_msqd,
                          color = "# of vessels landing squid (c/ landing sardine)"), size = 1) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = perc*coeff,
                          color = "% of vessels landing squid (c/ landings sardine)"),
            size = 1, linetype = "dashed") + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Number of vessels",
                     sec.axis = sec_axis(~./coeff, name = "% vessels harvesting squid" )) +
  theme(axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +  scale_color_manual(name = "Variables: ",
                     values = c("# of vessels landing sardine" = "green",
                                "# of vessels landing squid (c/ landing sardine)" = "blue",
                               "% of vessels landing squid (c/ landings sardine)" = "grey")) +
  theme(legend.position="right")

  rm(psdn.by.year, squid.cond.psdn, n_vessel_psdn, n_vessel_msqd_c_psdn, df, coeff)

msqd.by.year <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2020) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR) %>% filter(PACFIN_SPECIES_CODE  %in% c("MSQD")) %>%
  unique() %>% mutate(harv.msqd = 1)

# If a vessel harvest Pacific sardine in a year, how many also harvest squid.
psdn.cond.msqd <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2020) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR) %>% merge(msqd.by.year,
                        by = c("VESSEL_NUM", "LANDING_YEAR"), all.y = TRUE) %>%
  filter(harv.msqd == 1) %>% filter(PACFIN_SPECIES_CODE.x == "PSDN") %>% unique()

# Calculate number of vessels harvesting sardine and squid and percentage per year
n_vessel_msqd <- msqd.by.year %>% group_by(LANDING_YEAR) %>% summarize(n_vessel_msqd = sum(harv.msqd))
n_vessel_psdn_c_msqd <- psdn.cond.msqd %>% group_by(LANDING_YEAR) %>% summarize(n_vessel_psdn = sum(harv.msqd))
  df <- n_vessel_psdn_c_msqd %>%  mutate(n_vessel_msqd = n_vessel_msqd$n_vessel_msqd) %>%
    mutate(perc = n_vessel_psdn / n_vessel_msqd)

coeff = 200
g2 <- ggplot(df) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = n_vessel_msqd,
                          color = "# of vessels landing squid"), size = 1) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = n_vessel_psdn,
                          color = "# of vessels landing sardine (c/ landing squid)"), size = 1) +
  geom_line(mapping = aes(x = LANDING_YEAR, y = perc*coeff,
                          color = "% of vessels landing sardine (c/ landings squid)"),
            size = 1, linetype = "dashed") + scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Number of vessels",
                     sec.axis = sec_axis(~./coeff, name = "% vessels harvesting squid" )) +
  theme(axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +  scale_color_manual(name = "Variables: ",
                     values = c("# of vessels landing squid" = "green",
                                "# of vessels landing sardine (c/ landing squid)" = "blue",
                               "% of vessels landing sardine (c/ landings squid)" = "grey")) +
  theme(legend.position="right")

g1 / g2

rm(msqd.by.year, psdn.cond.msqd, n_vessel_msqd, n_vessel_psdn_c_msqd, df, coeff, g1, g2)

```

```{r switching_month, eval=FALSE, include=FALSE}
psdn.by.month <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR, LANDING_MONTH) %>% filter(PACFIN_SPECIES_CODE  %in% c("PSDN")) %>%
  unique() %>% mutate(harv.psdn = 1)

# If a vessel harvest Pacific sardine in a year, how many also harvest squid.
squid.cond.psdn <- PacFIN.month.CPS %>% filter(LANDING_YEAR >= 2000) %>% filter(LANDING_YEAR <= 2014) %>%
  select(VESSEL_NUM, PACFIN_SPECIES_CODE, LANDING_YEAR, LANDING_MONTH) %>% merge(psdn.by.month,
                        by = c("VESSEL_NUM", "LANDING_YEAR", "LANDING_MONTH"), all.y = TRUE) %>%
  filter(harv.psdn == 1) %>% filter(PACFIN_SPECIES_CODE.x == "MSQD") %>% unique()

# Calculate number of vessels harvesting sardine and squid and percentage per year
n_vessel_psdn <- psdn.by.month %>% group_by(LANDING_YEAR, LANDING_MONTH) %>%
  summarize(n_vessel_psdn = sum(harv.psdn, na.rm = TRUE))

n_vessel_msqd_c_psdn <- squid.cond.psdn %>% group_by(LANDING_YEAR, LANDING_MONTH) %>%
  summarize(n_vessel_msqd = sum(harv.psdn, na.rm = TRUE))

df <- merge(n_vessel_psdn, n_vessel_msqd_c_psdn, by=c("LANDING_YEAR", "LANDING_MONTH"), all.x = TRUE) %>%
    mutate(n_vessel_msqd = if_else(is.na(n_vessel_msqd), 0, n_vessel_msqd)) %>% group_by(LANDING_MONTH) %>%
    summarize(n_vessel_msqd = sum(n_vessel_msqd, na.rm = TRUE), n_vessel_psdn = sum(n_vessel_psdn, na.rm = TRUE)) %>%
    mutate(perc = (n_vessel_msqd / n_vessel_psdn)) %>% select(LANDING_MONTH, perc)

ggplot(data=df, aes(x=LANDING_MONTH, y=perc)) +
    geom_bar(stat="identity", fill=rgb(0.1,0.4,0.5,0.7), width=0.4) +
  scale_x_continuous(name = "Month", breaks=1:12) +
  scale_y_percent(name = "Percentage of vessels harvesting squid" )

rm(psdn.by.month, squid.cond.psdn, n_vessel_psdn, n_vessel_msqd_c_psdn, df)


```


# Methods


```{r dataset, include=FALSE}
# Organize dataset
PacFIN.month.dataset = PacFIN.month.cluster[,!grepl("*_LE_",names(PacFIN.month.cluster))]
PacFIN.month.dataset <- PacFIN.month.dataset %>% filter(LANDING_YEAR >= 2000) %>% # filter(PACFIN_SPECIES_CODE == "PSDN" | PACFIN_SPECIES_CODE == "MSQD") %>%
  dplyr::select(LANDING_YEAR, LANDING_MONTH, PORT_NAME, VESSEL_NUM, PSDN_SDM_60, MSQD_SDM_90,
                MSQD_SPAWN_SDM_90, LANDED_WEIGHT_MTONS.sum, AFI_PRICE_PER_KG.mean, 
                PACFIN_SPECIES_CODE, PACFIN_PORT_CODE, AGENCY_CODE, group) %>% 
  mutate(AFI_PRICE_PER_KG.mean = na_if(AFI_PRICE_PER_KG.mean, 0)) %>% 
  filter(group != is.na(group)) %>% 
  melt(id.vars=c("LANDING_YEAR", "LANDING_MONTH", "VESSEL_NUM",  "PORT_NAME", 
                 "PACFIN_PORT_CODE", "PACFIN_SPECIES_CODE", "AGENCY_CODE", "group")) %>% 
  dcast(LANDING_YEAR + LANDING_MONTH + VESSEL_NUM + PORT_NAME + PACFIN_PORT_CODE + AGENCY_CODE + group ~
          PACFIN_SPECIES_CODE + variable, fun.aggregate=mean, rm.na = T) %>%
  dplyr::select('LANDING_YEAR', 'LANDING_MONTH', 'VESSEL_NUM', 'PORT_NAME', 'PACFIN_PORT_CODE', 
         'AGENCY_CODE', 'group', 'PSDN_LANDED_WEIGHT_MTONS.sum', 'MSQD_LANDED_WEIGHT_MTONS.sum', 
         'PSDN_AFI_PRICE_PER_KG.mean', 'MSQD_AFI_PRICE_PER_KG.mean')

# Add port area code #  
PacFIN.month.dataset[PacFIN.month.dataset == "NaN"] <- NA
    port_area <- read.csv(file = here::here("Data", "Ports", "ports_area_and_name_codes.csv"))
    PacFIN.month.dataset <- PacFIN.month.dataset %>% 
      merge(port_area, by = c("PACFIN_PORT_CODE"), all.x = TRUE)
    
# Exclude port areas without landings of MSQD or PSDN #
selected.ports <- PacFIN.month.dataset %>% 
  group_by(PORT_AREA_CODE) %>% 
  summarise(landings_PSDN = sum(PSDN_LANDED_WEIGHT_MTONS.sum, na.rm = TRUE),
            landings_MSQD = sum(MSQD_LANDED_WEIGHT_MTONS.sum, na.rm = TRUE)) %>%
  rowwise() %>% mutate(total = min(landings_PSDN, landings_MSQD)) %>%
  filter(total > 0) %>% select(PORT_AREA_CODE) %>% mutate(port_included = 1)
  PacFIN.month.dataset <- PacFIN.month.dataset %>% 
    merge(selected.ports, by = c("PORT_AREA_CODE"), all.x = TRUE) %>%
    filter(port_included == 1) 

# Create ID data and change NaN to NA #
  
PacFIN.month.dataset <- PacFIN.month.dataset %>%
  mutate(PORT_ID = as.numeric(as.factor(PORT_NAME))) %>%
  mutate(PORT_CODE_ID = as.numeric(as.factor(PACFIN_PORT_CODE))) %>% 
  mutate(PORT_AREA_ID = as.numeric(as.factor(PORT_AREA_CODE))) 
  PacFIN.month.dataset[PacFIN.month.dataset == "NaN"] <- NA
  
# Merge SDM and ACL by year/month/port

## Merge data with SDM Pacific Sardine
SDM_port_PSDN <- read.csv(file = here::here("Data", "SDM", "PSDN_SDM_port_month.csv"))
  PacFIN.month.dataset <- merge(PacFIN.month.dataset, SDM_port_PSDN, 
                        by = c("PORT_NAME", "AGENCY_CODE","LANDING_YEAR", "LANDING_MONTH"), all.x = TRUE)

## Merge data with SDM Market Squid
SDM_port_MSQD <- read.csv(file = here::here("Data", "SDM", "MSQD_SDM_port_month.csv"))
  PacFIN.month.dataset <- merge(PacFIN.month.dataset, SDM_port_MSQD, 
                        by = c("PORT_NAME", "LANDING_YEAR", "LANDING_MONTH"), all.x = TRUE)

## Merge data with SDM Market Squid 
SDM_port_MSQD_Spawn <- read.csv(file = here::here("Data", "SDM", "MSQD_Spawn_SDM_port_month.csv"))
  PacFIN.month.dataset <- merge(PacFIN.month.dataset, SDM_port_MSQD_Spawn, 
                        by = c("PORT_NAME", "LANDING_YEAR", "LANDING_MONTH"), all.x = TRUE)  
  
dataset <- PacFIN.month.dataset %>%
    dplyr::rename(PSDN_SDM_60 = SDM_60) %>% dplyr::rename(MSQD_SDM_90 = SDM_90) %>%
    dplyr::rename(MSQD_SPAWN_SDM_90 = SDM_SPAWN_90)

  
rm(PacFIN.month.dataset, port_area, selected.ports, SDM_port_PSDN, SDM_port_MSQD, SDM_port_MSQD_Spawn)

```


```{r calculate-sdm, include=FALSE}
# Calculate year SDM for N/A's using PacFIN port code and port area code.

## Pacific sardine

### (a) Using PacFIN port code
SDM.port.code.PSDN <- aggregate(x=dataset$PSDN_SDM_60, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PACFIN_PORT_CODE), FUN = mean, na.rm=T)
  SDM.port.code.PSDN <- SDM.port.code.PSDN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PACFIN_PORT_CODE = Group.3) %>%
    dplyr::rename(PSDN.SDM.port.code = x)
    SDM.port.code.PSDN[SDM.port.code.PSDN == "NaN"] <- NA
    dataset <- dataset %>%
      merge(SDM.port.code.PSDN, 
            by = c("LANDING_YEAR", "LANDING_MONTH", "PACFIN_PORT_CODE"), all.x = TRUE) %>% 
      mutate(PSDN_SDM_60 = ifelse(is.na(PSDN_SDM_60), PSDN.SDM.port.code, PSDN_SDM_60))
        rm(SDM.port.code.PSDN)
    
### (b) Using port area code
SDM.port.area.PSDN <- aggregate(x=dataset$PSDN_SDM_60, by = list(dataset$LANDING_YEAR,
  dataset$LANDING_MONTH, dataset$PORT_AREA_CODE), FUN = mean, na.rm=T)
  SDM.port.area.PSDN <- SDM.port.area.PSDN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PORT_AREA_CODE = Group.3) %>%
    dplyr::rename(PSDN.SDM.port.area = x)
    SDM.port.area.PSDN[SDM.port.area.PSDN == "NaN"] <- NA
    dataset <- dataset %>%
      merge(SDM.port.area.PSDN, by = c("LANDING_YEAR", "LANDING_MONTH", "PORT_AREA_CODE"), all.x = TRUE) %>%
      mutate(PSDN_SDM_60 = ifelse(is.na(PSDN_SDM_60), PSDN.SDM.port.area, PSDN_SDM_60))
        rm(SDM.port.area.PSDN)

dataset = subset(dataset, select = -c(PSDN.SDM.port.code, PSDN.SDM.port.area))

## Market squid 

### (a) Using PacFIN port code
SDM.port.code.MSQD <- aggregate(x=dataset$MSQD_SDM_90, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PACFIN_PORT_CODE), FUN = mean, na.rm=T)
  SDM.port.code.MSQD <- SDM.port.code.MSQD %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PACFIN_PORT_CODE = Group.3) %>%
    dplyr::rename(MSQD.SDM.port.code = x)
    SDM.port.code.MSQD[SDM.port.code.MSQD == "NaN"] <- NA
    dataset <- dataset %>%
      merge(SDM.port.code.MSQD, 
            by = c("LANDING_YEAR", "LANDING_MONTH", "PACFIN_PORT_CODE"), all.x = TRUE) %>% 
      mutate(MSQD_SDM_90 = ifelse(is.na(MSQD_SDM_90), MSQD.SDM.port.code, MSQD_SDM_90))
        rm(SDM.port.code.MSQD)
    
### (b) Using port area code
SDM.port.area.MSQD <- aggregate(x=dataset$MSQD_SDM_90, by = list(dataset$LANDING_YEAR,
  dataset$LANDING_MONTH, dataset$PORT_AREA_CODE), FUN = mean, na.rm=T)
  SDM.port.area.MSQD <- SDM.port.area.MSQD %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PORT_AREA_CODE = Group.3) %>%
    dplyr::rename(MSQD.SDM.port.area = x)
    SDM.port.area.MSQD[SDM.port.area.MSQD == "NaN"] <- NA
    dataset <- dataset %>%
      merge(SDM.port.area.MSQD, by = c("LANDING_YEAR", "LANDING_MONTH", "PORT_AREA_CODE"), all.x = TRUE) %>%
      mutate(MSQD_SDM_90 = ifelse(is.na(MSQD_SDM_90), MSQD.SDM.port.area, MSQD_SDM_90))
        rm(SDM.port.area.MSQD)

dataset = subset(dataset, select = -c(MSQD.SDM.port.code, MSQD.SDM.port.area))


# Market squid (SPAWN)
## (a) Using PacFIN port code
 SDM.port.code.MSQD_SPAWN <- aggregate(x=dataset$MSQD_SPAWN_SDM_90, by = list(dataset$LANDING_YEAR,
   dataset$LANDING_MONTH, dataset$PACFIN_PORT_CODE), FUN = mean, na.rm=T)
   SDM.port.code.MSQD_SPAWN <- SDM.port.code.MSQD_SPAWN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
     dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PACFIN_PORT_CODE = Group.3) %>%
     dplyr::rename(MSQD_SPAWN.SDM.port.code = x)
     SDM.port.code.MSQD_SPAWN[SDM.port.code.MSQD_SPAWN == "NaN"] <- NA
     dataset <- dataset %>%
       merge(SDM.port.code.MSQD_SPAWN,
             by = c("LANDING_YEAR", "LANDING_MONTH", "PACFIN_PORT_CODE"), all.x = TRUE) %>%
       mutate(MSQD_SPAWN_SDM_90 = ifelse(is.na(MSQD_SPAWN_SDM_90), MSQD_SPAWN.SDM.port.code, MSQD_SPAWN_SDM_90))
         rm(SDM.port.code.MSQD_SPAWN)
 ### (b) Using port area code
 SDM.port.area.MSQD_SPAWN <- aggregate(x=dataset$MSQD_SPAWN_SDM_90, by = list(dataset$LANDING_YEAR,
   dataset$LANDING_MONTH, dataset$PORT_AREA_CODE), FUN = mean, na.rm=T)
   SDM.port.area.MSQD_SPAWN <- SDM.port.area.MSQD_SPAWN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
     dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PORT_AREA_CODE = Group.3) %>%
     dplyr::rename(MSQD_SPAWN.SDM.port.area = x)
     SDM.port.area.MSQD_SPAWN[SDM.port.area.MSQD_SPAWN == "NaN"] <- NA
     dataset <- dataset %>%
       merge(SDM.port.area.MSQD_SPAWN, by = c("LANDING_YEAR", "LANDING_MONTH", "PORT_AREA_CODE"), all.x = TRUE) %>%
       mutate(MSQD_SPAWN_SDM_90 = ifelse(is.na(MSQD_SPAWN_SDM_90), MSQD_SPAWN.SDM.port.area, MSQD_SPAWN_SDM_90))
         rm(SDM.port.area.MSQD_SPAWN)
 dataset = subset(dataset, select = -c(MSQD_SPAWN.SDM.port.code, MSQD_SPAWN.SDM.port.area))
     
```


Besides the biological stock, landings are affected by economic conditions. Some of them are harvest cost, prices received by species and their substitutes, and regulations imposed by the authorities. Our dataset includes <!-- average distances traveled by vessel from the port of origin *(TO BE INCLUDED USING GLOBAL FISHING WATCH)* and fuel cost *(TO BE INCLUDED...)* as proxies of harvest cost. We also include  --> the average species' price that a vessel receives in a port during a quarter, obtained from the PacFIN landings dataset, and they enter the model as both absolute and relative values. When prices were missing, we replaced this value with the average species' price in the port for all vessels in the corresponding year. If we still have missing values, we use the average price during the month at the port code, then at the port area, then at the state, and then considering the whole U.S. west coast (excluding Alaska). <!-- We should also include weather variables. Bad weather should reduce fishers' decision to go to sea. Ir becomes risky, but also this is compared to revenue and how risk-averse is a fisher (see Lisa's work): "Weather and tides are major factors in fishing operations and timing of vessels transiting in and out of the Columbia River." SAFE --> We include the Annual Catch Limit (ACL) for the Pacific sardine model as an additional explanatory variable in the landing equation, $\mu_i$. ACL variable is the total quota allocated each year for this fishery. We obtained this information from the CPS Fisheries Management Plan [@CPSFMP2021] and the Stock Assessment and Fishery Evaluation for 2019 document [@SAFE2020]. <!-- ISAAC's comments: Is this the ACL or the remaining part of the ACL (accounting for the previous catch). R: In the annual case does not matter, but in a monthly model, I should consider this. --> Our database comprises the period between 2000 and 2018. We do not include ACL as an explanatory variable as this variable has no variation between years. To capture any change in fishers' behavior when the Pacific sardine fishery is closed, we include a binary variable called **Closure** that takes the value "1" when the Pacific sardine fishery is closed and takes the value "0" when the Pacific sardine fishery is open. Concerning the Pacific sardine probability of presence, this variable takes the value of zero when the fishery is closed. Thus, the effect of this variable is only relevant to capture potential substitution between targets when the Pacific sardine fishery is open. We show a summary statistics of our data set in Table \ref{tab:sum_stats}. Specifically $\mu_{it}$ is defined as the following: 
$$ \begin{aligned} \mu_{i,p,t}   =  \beta^0_{p,c} &  + \beta^1_{p,c} P(Precense.PSDN)_{p,t} +  \beta^2_{p,c} P\left(Precense.PSDN\right)_{p,t} * P\left(Precense.MSQD\right)_{p,t} \\ & + \beta^3_{p,c} Price.PSDN_{p,t} + \beta^4_{p,c} ACL_{t}. \end{aligned}$$

In the case of $\text{logit}(p_{ijt}) = \bf{X}\gamma_p$  we can model the probability $p_{ijm}$ that vessel $i$ participate in a particular fishery at port $p$ in quarter $t$ specifically as: 
$$ \begin{aligned} \text{logit}(p_{i,t}) = \gamma^0_{c,p} & + \gamma^1_{c,p} P(Precense.PSDN)_{i,t} + \gamma^2_{c,p} P\left(Precense.PSDN\right)_{i,t}*P\left(Precense.MSQD\right)_{i,t} \\ & + \gamma^3_{c,p} Closure_t.\end{aligned}$$

where $ExpectedCatch_{p,t} = \frac{\sum_{k=0}^KP(Precense.PSDN)_{i,t-k}}{K}$. <!-- + \beta_2  Closure + \beta_3 Mean.revenue_j --> <!-- & + \beta_5 HHI_i + \beta_6 Percent.revenue_j + \beta_7 LCG_i + \beta_8 LI_i + \beta_9 Year.fished. --> Note that the outside option is interpreted as "No fishing in the particular fishery," which means that the vessel is participating in another fishery or is no fishing. 




```{r calculate-price, include=FALSE}

# Calculate year price for N/A's using PacFIN port code, port area code, state and year/month.

## Pacific sardine
  
### (a) Using port name
price.port.name.PSDN <- aggregate(x=dataset$PSDN_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PORT_NAME), FUN = mean, na.rm=T)
  price.port.name.PSDN <- price.port.name.PSDN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PORT_NAME = Group.3) %>%
    dplyr::rename(PSDN.price.port.name = x)
    price.port.name.PSDN[price.port.name.PSDN == "NaN"] <- NA
    dataset <- dataset %>% 
      merge(price.port.name.PSDN, by = c("LANDING_YEAR", "LANDING_MONTH", "PORT_NAME"), all.x = TRUE) %>%
      mutate(PSDN_AFI_PRICE_PER_KG.mean = ifelse(is.na(PSDN_AFI_PRICE_PER_KG.mean), 
        PSDN.price.port.name, PSDN_AFI_PRICE_PER_KG.mean))
        rm(price.port.name.PSDN)
    
### (b) Using PacFIN port code
price.port.code.PSDN <- aggregate(x=dataset$PSDN_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PACFIN_PORT_CODE), FUN = mean, na.rm=T)
  price.port.code.PSDN <- price.port.code.PSDN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PACFIN_PORT_CODE = Group.3) %>%
    dplyr::rename(PSDN.price.port.code = x)
    price.port.code.PSDN[price.port.code.PSDN == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.port.code.PSDN, by = c("LANDING_YEAR", "LANDING_MONTH", "PACFIN_PORT_CODE"), all.x = TRUE) %>% 
      mutate(PSDN_AFI_PRICE_PER_KG.mean = ifelse(is.na(PSDN_AFI_PRICE_PER_KG.mean), 
        PSDN.price.port.code, PSDN_AFI_PRICE_PER_KG.mean))
        rm(price.port.code.PSDN)
    
### (c) Using port area code
price.port.area.PSDN <- aggregate(x=dataset$PSDN_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PORT_AREA_CODE), FUN = mean, na.rm=T)
  price.port.area.PSDN <- price.port.area.PSDN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PORT_AREA_CODE = Group.3) %>%
    dplyr::rename(PSDN.price.port.area = x)
    price.port.area.PSDN[price.port.area.PSDN == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.port.area.PSDN, by = c("LANDING_YEAR", "LANDING_MONTH", "PORT_AREA_CODE"), all.x = TRUE) %>% 
      mutate(PSDN_AFI_PRICE_PER_KG.mean = ifelse(is.na(PSDN_AFI_PRICE_PER_KG.mean), 
        PSDN.price.port.area, PSDN_AFI_PRICE_PER_KG.mean))
        rm(price.port.area.PSDN)
        
### (d) Using state
price.state.PSDN <- aggregate(x=dataset$PSDN_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$AGENCY_CODE), FUN = mean, na.rm=T)
  price.state.PSDN <- price.state.PSDN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(AGENCY_CODE = Group.3) %>%
    dplyr::rename(PSDN.price.state = x)
    price.state.PSDN[price.state.PSDN == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.state.PSDN, by = c("LANDING_YEAR", "LANDING_MONTH", "AGENCY_CODE"), all.x = TRUE) %>% 
      mutate(PSDN_AFI_PRICE_PER_KG.mean = ifelse(is.na(PSDN_AFI_PRICE_PER_KG.mean), 
        PSDN.price.state, PSDN_AFI_PRICE_PER_KG.mean))
        rm(price.state.PSDN)
        
### (e) Using year/month
price.year.PSDN <- aggregate(x=dataset$PSDN_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH), FUN = mean, na.rm=T)
  price.year.PSDN <- price.year.PSDN %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PSDN.price.year.month = x)
    price.year.PSDN[price.year.PSDN == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.year.PSDN, by = c("LANDING_YEAR", "LANDING_MONTH"), all.x = TRUE) %>% 
      mutate(PSDN_AFI_PRICE_PER_KG.mean = ifelse(is.na(PSDN_AFI_PRICE_PER_KG.mean), 
        PSDN.price.year.month, PSDN_AFI_PRICE_PER_KG.mean))
        rm(price.year.PSDN)
    
dataset = subset(dataset, select = 
                   -c(PSDN.price.port.name, PSDN.price.port.code, 
                      PSDN.price.port.area, PSDN.price.state, PSDN.price.year.month))


## Market squid
  
### (a) Using port name
price.port.name.MSQD <- aggregate(x=dataset$MSQD_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PORT_NAME), FUN = mean, na.rm=T)
  price.port.name.MSQD <- price.port.name.MSQD %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PORT_NAME = Group.3) %>%
    dplyr::rename(MSQD.price.port.name = x)
    price.port.name.MSQD[price.port.name.MSQD == "NaN"] <- NA
    dataset <- dataset %>% 
      merge(price.port.name.MSQD, by = c("LANDING_YEAR", "LANDING_MONTH", "PORT_NAME"), all.x = TRUE) %>%
      mutate(MSQD_AFI_PRICE_PER_KG.mean = ifelse(is.na(MSQD_AFI_PRICE_PER_KG.mean), 
        MSQD.price.port.name, MSQD_AFI_PRICE_PER_KG.mean))
        rm(price.port.name.MSQD)
    
### (b) Using PacFIN port code
price.port.code.MSQD <- aggregate(x=dataset$MSQD_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PACFIN_PORT_CODE), FUN = mean, na.rm=T)
  price.port.code.MSQD <- price.port.code.MSQD %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PACFIN_PORT_CODE = Group.3) %>%
    dplyr::rename(MSQD.price.port.code = x)
    price.port.code.MSQD[price.port.code.MSQD == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.port.code.MSQD, by = c("LANDING_YEAR", "LANDING_MONTH", "PACFIN_PORT_CODE"), all.x = TRUE) %>% 
      mutate(MSQD_AFI_PRICE_PER_KG.mean = ifelse(is.na(MSQD_AFI_PRICE_PER_KG.mean), 
        MSQD.price.port.code, MSQD_AFI_PRICE_PER_KG.mean))
        rm(price.port.code.MSQD)
    
### (c) Using port area code
price.port.area.MSQD <- aggregate(x=dataset$MSQD_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$PORT_AREA_CODE), FUN = mean, na.rm=T)
  price.port.area.MSQD <- price.port.area.MSQD %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(PORT_AREA_CODE = Group.3) %>%
    dplyr::rename(MSQD.price.port.area = x)
    price.port.area.MSQD[price.port.area.MSQD == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.port.area.MSQD, by = c("LANDING_YEAR", "LANDING_MONTH", "PORT_AREA_CODE"), all.x = TRUE) %>% 
      mutate(MSQD_AFI_PRICE_PER_KG.mean = ifelse(is.na(MSQD_AFI_PRICE_PER_KG.mean), 
        MSQD.price.port.area, MSQD_AFI_PRICE_PER_KG.mean))
        rm(price.port.area.MSQD)
        
### (d) Using state
price.state.MSQD <- aggregate(x=dataset$MSQD_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH, dataset$AGENCY_CODE), FUN = mean, na.rm=T)
  price.state.MSQD <- price.state.MSQD %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(AGENCY_CODE = Group.3) %>%
    dplyr::rename(MSQD.price.state = x)
    price.state.MSQD[price.state.MSQD == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.state.MSQD, by = c("LANDING_YEAR", "LANDING_MONTH", "AGENCY_CODE"), all.x = TRUE) %>% 
      mutate(MSQD_AFI_PRICE_PER_KG.mean = ifelse(is.na(MSQD_AFI_PRICE_PER_KG.mean), 
        MSQD.price.state, MSQD_AFI_PRICE_PER_KG.mean))
        rm(price.state.MSQD)
        
### (e) Using year/month
price.year.MSQD <- aggregate(x=dataset$MSQD_AFI_PRICE_PER_KG.mean, by = list(dataset$LANDING_YEAR, 
  dataset$LANDING_MONTH), FUN = mean, na.rm=T)
  price.year.MSQD <- price.year.MSQD %>% dplyr::rename(LANDING_YEAR = Group.1) %>%
    dplyr::rename(LANDING_MONTH = Group.2) %>% dplyr::rename(MSQD.price.year.month = x)
    price.year.MSQD[price.year.MSQD == "NaN"] <- NA
    dataset <- dataset %>%
      merge(price.year.MSQD, by = c("LANDING_YEAR", "LANDING_MONTH"), all.x = TRUE) %>% 
      mutate(MSQD_AFI_PRICE_PER_KG.mean = ifelse(is.na(MSQD_AFI_PRICE_PER_KG.mean), 
        MSQD.price.year.month, MSQD_AFI_PRICE_PER_KG.mean))
        rm(price.year.MSQD)
    
dataset = subset(dataset, select = 
                   -c(MSQD.price.port.name, MSQD.price.port.code, 
                      MSQD.price.port.area, MSQD.price.state, MSQD.price.year.month))

```

```{r desc, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Rename Price and landing variables #
dataset <- dataset %>% 
  dplyr::rename(PSDN_Price = PSDN_AFI_PRICE_PER_KG.mean) %>%
  dplyr::rename(MSQD_Price = MSQD_AFI_PRICE_PER_KG.mean) %>%
  dplyr::rename(PSDN_Landings = PSDN_LANDED_WEIGHT_MTONS.sum) %>%
  dplyr::rename(MSQD_Landings = MSQD_LANDED_WEIGHT_MTONS.sum)

# Label dataset
sjlabelled::set_label(dataset$MSQD_SDM_90)         <- "Prob(presence): MSQD"
sjlabelled::set_label(dataset$MSQD_SPAWN_SDM_90)   <- "Prob(presence): MSQD (Spawn)"
sjlabelled::set_label(dataset$PSDN_SDM_60)         <- "Prob(presence): PSDN"
sjlabelled::set_label(dataset$LANDING_YEAR)     <- "Year"
sjlabelled::set_label(dataset$LANDING_MONTH)    <- "Month"
sjlabelled::set_label(dataset$PACFIN_PORT_CODE) <- "Port code"
sjlabelled::set_label(dataset$PORT_AREA_CODE)   <- "Port area code"
sjlabelled::set_label(dataset$PORT_NAME)        <- "Port name"
sjlabelled::set_label(dataset$AGENCY_CODE)      <- "State"
sjlabelled::set_label(dataset$VESSEL_NUM)       <- "Vessel ID"
sjlabelled::set_label(dataset$MSQD_Landings) <- "Landings: MSQD"
sjlabelled::set_label(dataset$MSQD_Price)    <- "Price: MSQD"
sjlabelled::set_label(dataset$PSDN_Landings) <- "Landings: PSDN"
sjlabelled::set_label(dataset$PSDN_Price)    <- "Price: PSDN"
sjlabelled::set_label(dataset$PORT_ID)       <- "Port ID"


### Convert to quarterly??? ###
library("zoo")
dataset$MonthYear <- as.yearmon(paste(dataset$LANDING_YEAR, dataset$LANDING_MONTH), "%Y %m")
dataset$QuarterYear <- as.yearqtr(dataset$MonthYear, format = "%Y-%m")

dataset <- dataset %>% 
  group_by(LANDING_YEAR, VESSEL_NUM, group, PORT_AREA_ID) %>%
  summarise(PSDN_SDM_60 = mean(PSDN_SDM_60, na.rm=T), MSQD_SDM_90 = mean(MSQD_SDM_90, na.rm=T),
            MSQD_SPAWN_SDM_90 = mean(MSQD_SPAWN_SDM_90, na.rm=T), PSDN_Landings = sum(PSDN_Landings, na.rm=T), 
            MSQD_Landings = sum(MSQD_Landings, na.rm=T), PSDN_Price = mean(PSDN_Price, na.rm=T), 
            MSQD_Price = mean(MSQD_Price, na.rm=T))
    dataset[dataset == "NaN"] <- NA

desc_data <- dataset %>%
  subset(select = -c(PORT_AREA_ID, VESSEL_NUM, group))

sjlabelled::set_label(desc_data$MSQD_SDM_90)       <- "Prob(presence): MSQD"
sjlabelled::set_label(desc_data$MSQD_SPAWN_SDM_90) <- "Prob(presence): MSQD (Spawn)"
sjlabelled::set_label(desc_data$PSDN_SDM_60)   <- "Prob(presence): PSDN"
sjlabelled::set_label(desc_data$MSQD_Landings) <- "Landings: MSQD"
sjlabelled::set_label(desc_data$MSQD_Price)    <- "Price: MSQD"
sjlabelled::set_label(desc_data$PSDN_Landings) <- "Landings: PSDN"
sjlabelled::set_label(desc_data$PSDN_Price)    <- "Price: PSDN"
sjlabelled::set_label(desc_data$ACL_PSDN)      <- "Annual Catch Limit: PSDN"


# Compute descriptive statistics #

desc_data <- desc_data[c("PSDN_SDM_60", "MSQD_SDM_90", "MSQD_SPAWN_SDM_90", "PSDN_Landings", "MSQD_Landings",
                         "PSDN_Price", "MSQD_Price")]
                                 
vtable::st(desc_data, labels = TRUE, title='Summary Statistics \\label{tab:sum_stats}', out='latex')
rm(desc_data, Tot.landings.ports)

```

<!-- ############################################### -->
<!-- ### CREATE DATASET BY SPECIES AND RUN MODEL ### -->
<!-- ############################################### -->

<!-- ### Pacific Sardine ### -->

```{r dataset_psdn, message=FALSE, warning=FALSE, include=FALSE, results='asis'}

## Select data for estimation, replace N/A landings to zero #

dataset_psdn <- dataset %>% 
  dplyr::select(VESSEL_NUM, PSDN_SDM_60, PSDN_Landings, PSDN_Price, MSQD_Price,
                PORT_AREA_ID, LANDING_YEAR, MSQD_SDM_90, MSQD_SPAWN_SDM_90, group) %>%
  filter(LANDING_YEAR >= 2000) %>% 
  dplyr::mutate(PSDN_Landings = coalesce(PSDN_Landings, 0)) %>% #filter(PSDN_Landings > 0) %>% 
  mutate(PSDN_Landings = ifelse(PSDN_Landings<=0, 0 ,PSDN_Landings)) %>% 
  mutate(Closure = ifelse(LANDING_YEAR >= 2015,1,0)) %>%
  mutate(RelPrice = PSDN_Price / MSQD_Price) %>% drop_na() 
  dataset_psdn$port_ID <- as.factor(
    udpipe::unique_identifier(dataset_psdn, fields = "PORT_AREA_ID", start_from = 1))
  dataset_psdn$cluster <- as.factor(
    udpipe::unique_identifier(dataset_psdn, fields = "group", start_from = 1))
# dataset_psdn_port_names <- dataset_psdn %>% dplyr::select(PORT_AREA_CODE, port_ID) %>% unique()

```

```{r psdn_estimation, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
fit_qPSDN <- brm(bf(PSDN_Landings ~ PSDN_SDM_60 + MSQD_SPAWN_SDM_90 + (1 | cluster) + (1 | port_ID), 
                    hu ~ (1 | cluster) + (1 | port_ID)),
                 data = dataset_psdn, 
                 prior = c(set_prior("cauchy(0,2)", class = "sd")),
                 family = hurdle_gamma(), 
                 chains = 4, cores = 4, warmup = "1000", iter = "2000",
                 control = list(adapt_delta = 0.95))

plot(fit_qPSDN)
plot(fit_qPSDN, pars = c("PSDN_SDM_60"))
coef(fit_qPSDN)

save.image (file = "stan_fit.RData")

# # work better with SDM separated. 
# loo(fit_qPSDN, fit_qPSDNv2)

```

```{r pp_check_psdn, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
g1 <- pp_check(fit_qPSDN) + ggtitle('(a) Pacific sardine') + 
  scale_color_manual(name = "", values = c("y" = "royalblue4", "yrep" = "azure3"),
                       labels = c("y" = "Observed", "yrep" = "Replicated")) +  
  theme(legend.position = "none", plot.title = element_text(size=9, face="bold.italic")) +  
  xlim(0.1, 4000) + ylab("Density")

g1
```



<!-- # print summary -->
<!-- texreg(list(f2, f3, r2, r3), caption = 'Panel data models for Pacific Sardine landings.\\label{table:sardine_est}', caption.above = TRUE, float.pos = "h", custom.model.names = c("FE: Model 1", "FE: Model 2", "RE: Model 1", "RE: Model 2")) -->


<!-- # ##  Extract Grouo-Level estimates ## -->
<!-- # ranef(fit_qPSDN_gamma) -->
<!-- #  -->
<!-- # ## Check divergence and other model check ## -->
<!-- # shinystan::launch_shinystan(fit_qPSDN_gamma) -->

<!-- # Investigate chain and posterior distributions.  -->
<!-- # https://cran.r-project.org/web/packages/bayesplot/vignettes/graphical-ppcs.html -->
<!-- # plot(fit_qPSDN_t2nc, pars = c("PSDN_SDM")) -->

<!-- # ### Predictions ### -->
<!-- # pred_data <- data.frame(PSDN_SDM = c(0.5, 0.25), MSQD_SDM = c(0.5), Port_ID = 1) -->
<!-- # predict(fit_qPSDN, newdata = pred_data, re_formula = NA) -->

<!-- # ## Compare models ## -->
<!-- # loo(fit1, fit2) -->


<!-- ### Market squid ### -->

```{r msqd_data, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
# Select data for estimation, replace N/A landings to zero #
est_data_msqd <- est_data %>%
  dplyr::select(port_num, Port, MSQD_SDM_90, MSQD_Landings, MSQD_Price, 
                  PSDN_SDM_60, NANC_SDM_20, Landing_year) %>%
  dplyr::mutate(MSQD_Landings = coalesce(MSQD_Landings, 0)) %>% 
  dplyr::mutate(dClose = ifelse(Landing_year >= 2015,1,0)) %>%
  dplyr::mutate(dOpen = ifelse(Landing_year < 2015,1,0)) %>%
  dplyr::mutate(PSDN_SDM_60_dOpen = PSDN_SDM_60 * dOpen) %>%
  filter(Landing_year >= 2000)

# Select ports that at least they have landing PSDN once, and drop N/A #
Tot.landings.ports <- as.data.frame(cbind(rowsum(est_data_msqd$MSQD_Landings, est_data_msqd$Port, na.rm = TRUE)))

Tot.landings.ports <- Tot.landings.ports %>%
  mutate(dTotalMSQD = ifelse(V1>0, 1, 0)) %>%
  mutate(Port = rownames(Tot.landings.ports)) %>%
  dplyr::select(Port, dTotalMSQD)

### NOTE: If I use drop_na, some ports without MSQD landings are excluded. ###
est_data_msqd <- est_data_msqd %>%
  merge(Tot.landings.ports, by.x = "Port", by.y = "Port", all.x = TRUE, all.y = FALSE) %>%
  filter(dTotalMSQD==1) %>%
  drop_na() 
  
# Create matrices and vectors to run STAN model #
year_id <- as.vector(cbind(as.numeric(factor(est_data_msqd$Landing_year))));
est_data_msqd$port_ID <- udpipe::unique_identifier(est_data_msqd, fields = "port_num", start_from = 1) 
portID_names_MSQD <- est_data_msqd %>%
  dplyr::select(Port, port_ID) %>%
  unique()

### Estimate using BRMS package ###
est_data_msqd$dClose    <- factor(est_data_msqd$dClose)
est_data_msqd$port_ID   <- factor(est_data_msqd$port_ID)
class(est_data_msqd$dClose)
class(est_data_msqd$port_ID)
class(est_data_msqd$dOpen)
```


```{r msqd_est_price, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
fit_qMSQD_price <- brm(bf(MSQD_Landings ~ MSQD_SDM_90 + MSQD_Price + PSDN_SDM_60_dOpen + NANC_SDM_20 + dClose 
                      + (1 + MSQD_SDM_90 + MSQD_Price + PSDN_SDM_60_dOpen + NANC_SDM_20 + dClose | port_ID),
                               hu ~ MSQD_SDM_90 + MSQD_Price + PSDN_SDM_60_dOpen + NANC_SDM_20 + dClose 
                      + (1 + MSQD_SDM_90 + MSQD_Price + PSDN_SDM_60_dOpen + NANC_SDM_20 + dClose | port_ID)),
                       data = est_data_msqd,
                       family = hurdle_gamma(),
                       control = list(adapt_delta = 0.999, max_treedepth = 20),
                       chains = 4, cores = 4)
save.image (file = "stan_fit.RData")
```

```{r pp_check_msqd, eval=FALSE, include=FALSE}
g2 <- pp_check(fit_qMSQD_price) + ggtitle('(b) Market Squid') +
  scale_color_manual(name = "", values = c("y" = "royalblue4", "yrep" = "azure3"),
                       labels = c("y" = "Observed", "yrep" = "Replicated")) + 
  theme(legend.position = "none", plot.title = element_text(size=9, face="bold.italic"))  + 
  xlim(0.1, 15000) + xlab("Landing (tons)")

```

 


# Results

## Landing model


### Graphical posterior predictive
Before presenting the results for the three species in consideration, we check graphically whether the posterior distribution is able to predict the actual distribution. We exclude zero in our graph to avoid plotting different data generation processes. In Figure \ref{fig:posteriors} we show a posterior sample compare with the true sample distribution for the three species in analysis. Some deviation from the true sample distribution is observed in the three models, but still the posterior sample follow a similar shape than the actual curve. Moreover, our pacific sardine sample do well in predicting probabilities at lower landing levels. 

```{r y_rep, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
g1 

```

In addition to a posterior predictive check, we also check divergence and treedepth.^[Explain what are these concepts...] In all three models, we did not have any divergent step in our estimation. The maximum treedepth was 11 for the Pacific sardine model, while in the others it was 10.   

```{r shinystan, eval=FALSE, include=FALSE}
shinystan::launch_shinystan(fit_qPSDN_price) 
shinystan::launch_shinystan(fit_qMSQD_price) 
shinystan::launch_shinystan(fit_qNANC_price) 
```

```{r y_rep_zero, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
g1 <- pp_check(fit_qPSDN_price) + ggtitle('(a) Pacific sardine')  + theme(legend.position = "none") + xlim(0, 0.1) 
g2 <- pp_check(fit_qMSQD_price) + ggtitle('(b) Market Squid')     + theme(legend.position = "none") + xlim(0, 0.1) 
g3 <- pp_check(fit_qNANC_price) + ggtitle('(c) Northern anchovy') + theme(legend.position = "right", 
                                             plot.title = element_text(size=9, face="bold.italic")) + xlim(0, 0.1) 



g1 + g2 + g3
```


```{r model-comparision, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
fit_qMSQD <- add_criterion(fit_qMSQD, "loo")
fit_qMSQD_90 <- add_criterion(fit_qMSQD_90, "loo")

comp <- loo_compare(loo(fit_qMSQD), loo(fit_qMSQD_90)) %>%
  as.data.frame() %>%
  select(elpd_loo, elpd_diff, se_diff, p_loo, looic) %>%
  dplyr::rename(
    "ELPD-Diff" = elpd_diff, "SE-Diff" = se_diff, 
    "ELPD-LOO" = elpd_loo, "P-LOO" = p_loo, "LOOIC" = looic
  )

mw1 <- model_weights(fit_qMSQD, fit_qMSQD_90, weights = "loo")[rownames(comp)]

comp %>%
  cbind("Akaike-Weight" = mw1) %>%
  apa_table(
    format = "latex", booktabs = TRUE, digits = 2,
    caption = "Comparison of models fit1 to fit3 based on approximate leave-one-out cross-validation. Market squid landigns model.",
    note =  "ELPD-LOO = expected log posterior predictive density (higher is better); ELPD-DIFF = difference in ELPD values compared to the best model. SE-DIFF = standard error of the ELPD difference. P-LOO = effective number of model parameters (lower is better); LOOIC: leave-one-out information criterion (lower is better); Akaike-Weight = Model weight based on the LOOIC values (higher is better).",
    align = c("l", rep("r", 6))
  )
```


### Own species distribution effect

The inclusion of hierarchical effects by port allows us to analyze the effects of the probability of presence on landings disaggregated by ports. We only focus our analysis in owns probability of presence effects, as other probability are analyzed as interaction (see next subsection). We can observe in Figure \ref{fig:sdmeffects} that the effects of the own probability of presence on landings are not strong (maybe I should use logs...). However, it is clear that it have a positive effect on landings, specially on port areas such as Columbia River at Oregon for Pacific sardine, and Monterrey for Northern anchovy, while a more moderate effect is observed at Santa Barbara Area for Northern anchovy.


```{r by_port_sdm, eval=FALSE, fig.cap=, include=FALSE}
# PSDN plots
conditions <- data.frame(port_ID = unique(est_data_psdn$port_ID))
  rownames(conditions) <- unique(est_data_psdn$Port)
conditions_psdn <- conditions %>% 
  rownames_to_column('port_name') %>%
  filter(port_ID == 1 | port_ID == 2 | port_ID == 3) %>%
  column_to_rownames('port_name')

c_eff_psdn <- (conditional_effects
               (fit_qPSDN_price, "PSDN_SDM_60", surface=TRUE, conditions = conditions_psdn, re_formula = NULL))
               #, transform = log, method = "posterior_predict"))
g1 <- plot(c_eff_psdn, plot = FALSE, nrow = 3, ncol = 1)[[1]] + ggtitle('(a) Pacific sardine')+ 
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
  scale_x_continuous(name = element_blank()) +
  scale_y_continuous(name = "Landings (tons)")

# MSQD plots
conditions2 <- data.frame(port_ID = unique(est_data_msqd$port_ID))
  rownames(conditions2) <- unique(est_data_msqd$Port)
conditions_msqd <- conditions2 %>% 
  rownames_to_column('port_name') %>%
  filter(port_ID == 4  | port_ID == 5  | port_ID == 8) %>%
  column_to_rownames('port_name')
c_eff_msqd <- (conditional_effects
               (fit_qMSQD_price, "MSQD_SDM_90", surface=TRUE, conditions = conditions_msqd, re_formula = NULL))
               #, transform = log, method = "posterior_predict"))
g2 <- plot(c_eff_msqd, plot = FALSE, nrow = 3, ncol = 1)[[1]] + ggtitle('(b) Market squid') + 
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
  scale_x_continuous(name = "Prob(Presence)") +
  scale_y_continuous(name = element_blank()) 

# NANC plots
conditions3 <- data.frame(port_ID = unique(est_data_nanc$port_ID))
  rownames(conditions3) <- unique(est_data_nanc$Port)
conditions_nanc <- conditions3 %>% 
  rownames_to_column('port_name') %>%
  filter(port_ID == 3  | port_ID == 4  | port_ID == 5) %>%
  column_to_rownames('port_name')
c_eff_nanc <- (conditional_effects
               (fit_qNANC_price, "NANC_SDM_20", surface=TRUE, conditions = conditions_nanc, re_formula = NULL))
               #, transform = log, method = "posterior_predict"))
g3 <- plot(c_eff_nanc, plot = FALSE, nrow = 3, ncol = 1)[[1]] + ggtitle('(c) Northern anchovy') + 
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
  scale_x_continuous(name = element_blank()) +
  scale_y_continuous(name = element_blank()) 

# Merge plots
g1 + g2 + g3
```

<!-- @smith2021potential estimate with GAM framework (allows for non-linear relationships) -->


### Interaction effects

To study the effect of other species distribution on landings, we compute the conditional effects of the interaction between probabilities of presence. Let denote species A the species which we are analysing its landings, and species B and C the other species included in the model for species A landings. For each species, we compute the effect on landing of the interaction between the probability of presences of: 
  + Species A and species B, 
  + Species A versus species C,
  + Species B versus species C.
For example, for the Pacific sardine equation, we compute these results for the probability of presence effect between Pacific sardine and market squid, Pacific sardine and Northern anchovy, and market squid and Northern anchovy. 

```{r int_effect_PSDN_by_port, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
c_eff_int_psdn_msqd <- (conditional_effects(fit_qPSDN_price, "PSDN_SDM_60:MSQD_SDM_90", surface=TRUE, 
                                            conditions = conditions_psdn, re_formula = NULL))
c_eff_int_psdn_nanc <- (conditional_effects(fit_qPSDN_price, "PSDN_SDM_60:NANC_SDM_20", surface=TRUE, 
                                            conditions = conditions_psdn, re_formula = NULL))
c_eff_int_msqd_nanc <- (conditional_effects(fit_qPSDN_price, "MSQD_SDM_90:NANC_SDM_20", surface=TRUE, 
                                            conditions = conditions_psdn, re_formula = NULL))

g1_PSDN <-  plot(c_eff_int_psdn_msqd, plot = FALSE)[[1]] + ggtitle('(a) Pacific sardine x Market squid') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: PSDN")) +
  scale_x_continuous(name = "P(Pres): PSDN") + scale_y_continuous(name = "P(Pres): MSQD") 

g2_PSDN <-  plot(c_eff_int_psdn_nanc, plot = FALSE)[[1]] + ggtitle('(b) Pacific sardine x Northern anchovy') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: PSDN")) + 
        scale_x_continuous(name = "P(Pres): PSDN") + scale_y_continuous(name = "P(Pres): NANC") 

g3_PSDN <-  plot(c_eff_int_msqd_nanc, plot = FALSE)[[1]] + ggtitle('(c) Northern anchovy x Market squid') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: PSDN")) + 
        scale_x_continuous(name = "P(Pres): MSQD") + scale_y_continuous(name = "P(Pres): NANC") 
save.image (file = "stan_fit.RData")
```

Let first analyze the interaction effect of the probability of presences on Pacific sardine landings, presented in Figure \ref{fig:sdm_int_psdn}. We only shows results for the main three port where Pacific sardine is landed (Columbia River at Oregon, Los Angeles Area  and Monterrey Area). As we already commented in Figure \ref{fig:sdmeffects}, the effects the Pacific sardine's own probability of presence is positive, changing color from left to right in panels (a) and (b). Market squid seems to be an stronger substitute for Pacific sardine than Northern anchovy. This is suggested by the different slopes between panel (a) and panel (b). For instane, in Columbia River, an increase in the probability of presence for anchovy does decrease Pacific sardine landings as much as an increase in the probability of presece of market squid. 

```{r int_effect_PSDN_by_port_PLOT, eval=FALSE, fig.cap=, message=FALSE, warning=FALSE, include=FALSE}
g1_PSDN / g2_PSDN
```

```{r int_effect_MSQD_by_port, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
c_eff_int_msqd_psdn <- (conditional_effects(fit_qMSQD_price, "MSQD_SDM_90:PSDN_SDM_60_dOpen", surface=TRUE, 
                                            conditions = conditions_msqd, re_formula = NULL))
c_eff_int_msqd_nanc <- (conditional_effects(fit_qMSQD_price, "MSQD_SDM_90:NANC_SDM_20", surface=TRUE, 
                                            conditions = conditions_msqd, re_formula = NULL))
c_eff_int_psdn_nanc <- (conditional_effects(fit_qMSQD_price, "PSDN_SDM_60_dOpen:NANC_SDM_20", surface=TRUE, 
                                            conditions = conditions_msqd, re_formula = NULL))

g1_MSQD <-  plot(c_eff_int_msqd_psdn, plot = FALSE)[[1]] + ggtitle('(a) Market squid x Pacific sardine') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: MSQD")) +
  scale_x_continuous(name = "P(Pres): MSQD") + scale_y_continuous(name = "P(Pres): PSDN") 

g2_MSQD <-  plot(c_eff_int_msqd_nanc, plot = FALSE)[[1]] + ggtitle('(b) Market squid x Northern anchovy') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: MSQD")) + 
        scale_x_continuous(name = "P(Pres): MSQD") + scale_y_continuous(name = "P(Pres): NANC") 

g3_MSQD <-  plot(c_eff_int_psdn_nanc, plot = FALSE)[[1]] + ggtitle('(c) Pacific sardine x Northern anchovy') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: MSQD")) + 
        scale_x_continuous(name = "P(Pres): PSDN") + scale_y_continuous(name = "P(Pres): NANC") 
save.image (file = "stan_fit.RData")
```

We found some similar findings for market squid in Figure \ref{fig:sdm_int_msqd}. Pacific sardine it is a strong substitute of market squid. In regard to Norther anchovy, there is some complementarity with market squid. We observe that when Northern anchovy presence increase, there is also an increase in market squid landings. 

```{r int_effect_MSQD_by_port_PLOT, eval=FALSE, fig.cap=, message=FALSE, warning=FALSE, include=FALSE}
g1_MSQD / g2_MSQD 
```


```{r int_effect_NANC_by_port, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
conditions_nanc <- conditions_nanc %>% filter(port_ID == 3  | port_ID == 5)

c_eff_int_nanc_psdn <- (conditional_effects(fit_qNANC_price, "NANC_SDM_20:PSDN_SDM_60_dOpen", surface=TRUE, 
                                            conditions = conditions_nanc, re_formula = NULL))
c_eff_int_nanc_msqd <- (conditional_effects(fit_qNANC_price, "NANC_SDM_20:MSQD_SDM_90", surface=TRUE, 
                                            conditions = conditions_nanc, re_formula = NULL))
c_eff_int_psdn_msqd <- (conditional_effects(fit_qNANC_price, "PSDN_SDM_60_dOpen:MSQD_SDM_90", surface=TRUE, 
                                            conditions = conditions_nanc, re_formula = NULL))

g1_NANC <-  plot(c_eff_int_nanc_psdn, plot = FALSE)[[1]] + ggtitle('(a) Northern anchovy x Pacific sardine') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: PSDN")) +
  scale_x_continuous(name = "P(Pres): NANC") + scale_y_continuous(name = "P(Pres): PSDN") 

g2_NANC <-  plot(c_eff_int_nanc_msqd, plot = FALSE)[[1]] + ggtitle('(b) Northern anchovy x Market squid') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: PSDN")) + 
        scale_x_continuous(name = "P(Pres): NANC") + scale_y_continuous(name = "P(Pres): MSQD") 

g3_NANC <-  plot(c_eff_int_psdn_msqd, plot = FALSE)[[1]] + ggtitle('(c) Pacific sardine x Market squid') + 
         theme(
          plot.title = element_text(size=9, face="bold.italic"), 
          axis.text = element_text(size = 7), 
          axis.title = element_text(size = 8),
          legend.title = element_text(size = 9), 
          legend.text = element_text(size=8)
        ) + guides(colour=guide_legend(title="Landings: PSDN")) + 
        scale_x_continuous(name = "P(Pres): PSDN") + scale_y_continuous(name = "P(Pres): MSQD") 
save.image (file = "stan_fit.RData")
```

<!-- ISAAC COMMENTS: There are some notes from two stakeholder workshops (Desiree et al.). It would be good to use these to understand and motivate the discussion of complementarity and substition.  -->

However, this complementary is not observed for Northern anchovy landings. When market squid presence increase, we observe a decrease in Northern anchovy landings. An interesting case is observed for Pacific sardine and Northern anchovy, in particular at Monterrey area. For small levels of Pacific sardine presence, Pacific sardine is a complement with Northern anchovy. However, after certain threshold, Pacific sardine is no longer complement and Pacific sardine become a substitute, reducing Northern anchovy landings when the probability of presence for Pacific sardine increase.

```{r int_effect_NANC_by_port_PLOT, eval=FALSE, fig.cap=, message=FALSE, warning=FALSE, include=FALSE}
g1_NANC / g2_NANC 
```


### Pacific sardine closure
Finally, lets analyze the effect of the Pacific sardine closure on market squid and Northern anchovy landings. We do not find significant results in the case of Northern anchovy. In the case of market squid, we found a decrease in the level of landing after the closure was implemented. It seems that some fisher exit the CPS fishery after the closure of Pacific sardine, reducing their participation in the market squid fishery. Our participation model results give us more details about what happen after the Pacific sardine closure.

```{r by_port_msqd_dclose, eval=FALSE, fig.cap=, include=FALSE}
# conditions_dClose <- data.frame(port_ID = unique(est_data_msqd$dClose))
# rownames(conditions_dClose) <- unique(est_data_msqd$dClose)
c_eff_close_msqd <- (conditional_effects(fit_qMSQD_price, "dClose", conditions = conditions_msqd, re_formula = NULL))
g1 <-  plot(c_eff_close_msqd, plot = FALSE)[[1]] + ggtitle('(a) Market squid') + 
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "Landings (tons)") 

c_eff_close_nanc <- (conditional_effects(fit_qNANC_price, "dClose", conditions = conditions_nanc, re_formula = NULL))
g2 <- plot(c_eff_close_nanc, plot = FALSE)[[1]] + ggtitle('(b) Northern anchovy') + 
  theme(plot.title = element_text(size=9, face="bold.italic"), 
        axis.text = element_text(size = 7), axis.title = element_text(size = 8)) + 
  scale_x_discrete(name = "Closure? (1 = True; 0 = False)") +
  scale_y_continuous(name = "Landings (tons)") 

g1 / g2
```

# Conclusions
A possible extension of our research is to consider spatial autocorrelations between ports.^[See \cite{morris2019bayesian} for an application of a spatial model in a Bayesian framework.] Ports landing maybe correlated as vessel have the incentives to choose the port of landing, conditional on whether the port have the infrastructure for this. It is likely that they just land wherever is closer to the area they are fishing. Nevertheless, differential in prices could encourage them to travel a little further for higher prices. 




<!-- ```{r int_effect_sep, eval=FALSE, include=FALSE} -->

<!-- hu <- plot(conditional_effects(fit_qPSDN, "PSDN_SDM:MSQD_SDM", surface=TRUE, dpar="hu"), -->
<!--            plot = FALSE, ask = FALSE) -->
<!-- mu <- plot(conditional_effects(fit_qPSDN, "PSDN_SDM:MSQD_SDM", surface=TRUE, dpar="mu"), -->
<!--            plot = FALSE, ask = FALSE) -->

<!-- mu[[1]] + hu[[1]] + plot_layout(nrow = 2) -->

<!-- ``` -->

<!-- ```{r int_effect_sep_msqd, eval=FALSE, fig.cap=, include=FALSE} -->

<!-- hu <- plot(conditional_effects(fit_qMSQD_gamma, "PSDN_SDM:MSQD_SDM", surface=TRUE, dpar="hu"), -->
<!--            plot = FALSE, ask = FALSE) -->
<!-- mu <- plot(conditional_effects(fit_qMSQD_gamma, "PSDN_SDM:MSQD_SDM", surface=TRUE, dpar="mu"), -->
<!--            plot = FALSE, ask = FALSE) -->

<!-- mu[[1]] + hu[[1]] + plot_layout(nrow = 2) -->

<!-- ``` -->


<!-- ## Effort susbsitution ##  -->

<!-- * Time series of number of trips (as a proxy for effort) for all the species -->
<!-- * @richerson2017 use a method identify the nature of the outliers in an ARMA time series model (read more if interested)  -->
<!--     + I propose to estimate a system of simultaneous equations (VECM model) to study equilibrium of effort and short-run and long-run effects of the closure (structural breaks) -->
<!--     + Have a long-run equation for each species (simultaneously estimated) and test for structural break in this long-run relationship. -->


<!-- ## Seasonality changes ## -->

<!-- * Seasonality can be studied calculating monthly share of total trips by species, regress it using month dummiues and see any is there any structural change after the closure [@richerson2017].  -->

<!-- <!-- Shall we study also effort as number of trips, and seasonality using time series -->


