---
title: "Future Seas "
subtitle: "Econ Report"
author: "Felipe Quezada"
institute: "UC Santa Cruz & NOAA SWFSC"
date: "September 2nd, 2021"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      countdown: 30000
---

<!-- xaringan:::inf_mr() -->

```{css, echo=FALSE}
# CSS for including pauses in printed PDF output (see bottom of lecture)
@media print {
  .has-continuation {
    display: block !important;
  }
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=4, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )
library(bookdown)
library(doBy)
library(dplyr)
library(fontawesome)
library(ggplot2)
library(hrbrthemes)
library(lmtest)
library(lubridate)
library(magrittr)
library(plm)
library(patchwork)
library(reshape)
library(reshape2)
library(scales)
library(sjlabelled)
library(summarytools)
library(texreg)
library(tidyr)
library(tidyverse)
library(tinytex)
library(viridis)
library(xtable)
library(zoo)
library(here) 
```
```{r data_import, include=FALSE}
PacFIN_dat <- read.csv(file = here::here("Data", "PacFin.csv"))

collapse <- PacFIN_dat %>%
  dplyr::filter(Species_code == "PSDN" | Species_code == "MSQD"  | Species_code == "NANC") %>%
  dplyr::select(Landing_year, Port, PSDN_SDM_60, MSQD_SDM_90, NANC_SDM_90, Landings, Species_code) %>% 
  dplyr::filter(Port == "LAA" | Port == "MNA" | Port == "SBA") %>%
      mutate(Port = fct_relevel(Port, "LAA", "SBA", "MNA")) %>% 
  spread(Species_code, Landings) %>% 
  dplyr::rename(Landings_PSDN = PSDN) %>% dplyr::rename(Landings_MSQD = MSQD) %>% dplyr::rename(Landings_NANC = NANC) %>%
  filter(Landing_year >= 2000 & Landing_year <= 2018)
```

```{r general_functions, include=FALSE}

meanfun <- function(x, ...){
  c(mean=mean(x, na.rm=TRUE, ...)) #, v=var(x, na.rm=TRUE, ...), l=length(x))
}

sumfun <- function(x, ...){
  c(sum=sum(x, na.rm=TRUE, ...)) #, v=var(x, na.rm=TRUE, ...), l=length(x))
}

sum_mean_fun <- function(x, ...){
  c(mean=mean(x, na.rm=TRUE, ...), sum=sum(x, na.rm=TRUE, ...)) #, l=length(x))
}

```

# Introduction

+ Research questions:
  + How do fishers decide between species?
  + How catch composition will be affected due to climate change?
  
+ Two approach to address this question:
  + Landing model for Pacific sardine, market squid and Northern anchovy.
  + Discrete choice model for location decision and species choice. 

---
class: inverse, center, middle
name: landingmodel

# Landing model

<html><div style='float:left'></div><hr color='#fdc700' size=1px width=796px></html>


---

# Landing model

+ Separate landing model for each species.

+ Interaction between species on landing
  + Probability of presence (SDMs outputs)
  
+ Multilevel Bayesian hurdle model $$[\mu_i, \sigma_{\mu}^2, p_i | q_{i,t}] \varpropto f\left(q_{i,t} | \mu_i, \sigma_{\mu}^2, p_i\right) \times [\mu_i][\sigma_{\mu}^2][p_i]$$
$$f\left(q_{i,t} | \mu_i, \sigma_{\mu}^2, p_i \right) = \begin{cases} p_i & \text{if} \quad q_{it} = 0  \\ \left[1-p_i\right] \text{gamma} \left(q_{i,t} | \frac{\mu_i^2}{\sigma_{\mu}^2}, \frac{\mu_i}{\sigma_{\mu}^2} \right) & \text{if} \quad q_{it} > 0. \end{cases}$$
---

# Results 

```{r int_effect_MSQD_by_port, echo=FALSE, message=FALSE, warning=FALSE}
plot(conditional_effects(
  fit_qPSDN_90, "PSDN_SDM_60:MSQD_SDM_90", surface=TRUE, conditions = conditions, 
  re_formula = NULL), ncol = 4, points = FALSE)
```

```{r int_effect_NANC_by_port, echo=FALSE, message=FALSE, warning=FALSE}
plot(conditional_effects(
  fit_qPSDN_90, "PSDN_SDM_60:NANC_SDM_90", surface=TRUE, conditions = conditions, 
  re_formula = NULL), ncol = 4, points = FALSE)
```

```{r by_port_both_sdm_msqd, echo=FALSE, message=FALSE, warning=FALSE}
conditions_dClose <- data.frame(port_ID = unique(est_data_msqd_all$dClose))
rownames(conditions_dClose) <- unique(est_data_msqd_all$dClose)

plot(conditional_effects(
  fit_qMSQD_90, "PSDN_SDM_60_dOpen:MSQD_SDM_90", surface=TRUE, conditions = conditions, 
  re_formula = NULL), ncol = 4, points = FALSE)
```

```{r by_port_both_sdm_nanc, echo=FALSE, message=FALSE, warning=FALSE}
conditions_dClose <- data.frame(port_ID = unique(est_data_msqd_all$dClose))
rownames(conditions_dClose) <- unique(est_data_msqd_all$dClose)

plot(conditional_effects(
  fit_qMSQD_90, "NANC_SDM_90:MSQD_SDM_90", surface=TRUE, conditions = conditions, 
  re_formula = NULL), ncol = 4, points = FALSE)
```
```{r by_port_msqd_dclose}
plot(conditional_effects(
  fit_qMSQD_90, "dClose", conditions = conditions, 
  re_formula = NULL), ncol = 4, points = FALSE)
```

---

class: inverse, center, middle
name: discrete-choice

# Discrete choice model

<html><div style='float:left'></div><hr color='#fdc700' size=1px width=796px></html>


---

# Discrete choice model

+ How species distribution affect fishers decisions?
    + Require logbook data for different species during same time period
        + Not available.
        + Merge landing data to Global Fishing Watch to recreate "logbooks"
    + Choice set:
        + Hicks (2020) to sample choice set from logbooks (thanks Peter for the code...)
    + SDM will be included to calculate exepcted catch. 
        + Forecasting
  
---

class: inverse, center, middle

# Thank you for your attention!

`r fa('address-book')` [felipequezada.com](https://felipequezada.com/)

`r fa('envelope')` [fequezad@ucsc.edu](mailto:fequezad@ucsc.edu)

`r fa('graduation-cap')` Postdoctoral Scholar, UC Santa Cruz & NOAA-SWFSC

<!-- ```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE} -->
<!--  infile = list.files(pattern = '.html') -->
<!--  pagedown::chrome_print(input = infile, timeout = 100) -->
<!-- ``` -->

